---
title: "Capstone Project - Week1-8"
author: "Amol Gote"
date: "12/18/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library_setup, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
library(tidyverse)
library(caret)
library(corrplot)
library(tidyverse)
library(ROCR)
library(gbm)
library(ROSE)
library(gridExtra)
library(parallel) 
library(tree)
library(rpart)
library(rpart.plot)
```


```{r load_data, MESSAGE = FALSE, warning=FALSE}
fundamentals_ds <- read.csv("./data/Fundamentals_DS.csv", na.strings=c(""," "))
nrow(fundamentals_ds)
```

### Filter the dataset with the sector assigned.
```{r load_data_for_specific_sector, MESSAGE = FALSE, warning=FALSE}
fundamentals_ds_sjm <- fundamentals_ds %>%
  filter(tic == 'SJM')
gsector_sjm <- head(fundamentals_ds_sjm$gsector, 1)
fundamentals_ds <- fundamentals_ds %>%
  filter(gsector == gsector_sjm)
nrow(fundamentals_ds)
gsector_sjm
```

### Filter the data to restrict the dataset to contain all annually reported statements, and exclude restatements.
```{r filter_data_datafmt, MESSAGE = FALSE, warning=FALSE}
names(fundamentals_ds)[names(fundamentals_ds) == "誰..gvkey"] <- "gvkey"
fundamentals_ds_filter <- fundamentals_ds %>%
  filter(datafmt == 'STD') 
nrow(fundamentals_ds_filter)

target_company <- fundamentals_ds_filter %>%
  filter(tic == 'SJM')
target_company_gv_key <- head(target_company$gvkey, 1)
target_company_gv_key
ncol(fundamentals_ds_filter)
```

```{r }
#summary(fundamentals_ds_filter)
```

````{r}
summary(fundamentals_ds_filter$mkvalt)
```

### Remove all the columns which contain 25% or greater NA or null values 
```{r remove_cols_with_high_na, MESSAGE = FALSE, warning=FALSE}
fundamentals_ds_filter_1 <- fundamentals_ds_filter[ lapply( fundamentals_ds_filter, function(x) sum(is.na(x)) / length(x) ) < 0.25 ]
ncol(fundamentals_ds_filter_1)
nrow(fundamentals_ds_filter_1)
```

````{r}
colnames(fundamentals_ds_filter_1)
```

```{r write_file, MESSAGE = FALSE, warning=FALSE}
write.csv(fundamentals_ds_filter_1, file = "data/fundamentals_ds_filter_1.csv", row.names=FALSE)
```

### Remove all the columns which are related to general information of the company like address, phone, url etc...
```{r write_filter_after_na_removal, MESSAGE = FALSE, warning=FALSE}
#fundamentals_ds_filter_1 <- fundamentals_ds_filter_1[ lapply( fundamentals_ds_filter_1, function(x) sum(is.na(x)) / length(x) ) < 0.25 ]
ncol(fundamentals_ds_filter_1)
fundamentals_ds_filter_1 <- subset(fundamentals_ds_filter_1, select = -c(datadate,indfmt,curncd,consol,popsrc,conm,curcd,apdedate,fdate,add1,addzip,busdesc,
                                                                         city,conml,weburl,phone,loc,final,fyr,acchg,fic,
                                                                         xido,xidoc,naicsh,sich,au,auop,auopic,fyrc,ggroup,gind,gsector,gsubind,priusa))
write.csv(fundamentals_ds_filter_1, file = "data/fundamentals_ds_filter_2.csv", row.names=FALSE, na="")
ncol(fundamentals_ds_filter_1)
nrow(fundamentals_ds_filter_1)
```

### Identify columns which have values which are new to zero variance
```{r identify_near_zero_variance, MESSAGE = FALSE, warning=FALSE}
nzv_ds <- nearZeroVar(fundamentals_ds_filter_1, saveMetrics = TRUE)
nzv_ds <- nzv_ds[nzv_ds[,"nzv"] > 0, ]
nzv_ds
```


### Remove columns which have values which are of zero variance
```{r remove_near_zero_variance_columns, MESSAGE = FALSE, warning=FALSE}
nzv_ds_cols <- nearZeroVar(fundamentals_ds_filter_1)
fundamentals_ds_filter_1 <- fundamentals_ds_filter_1[, -nzv_ds_cols]
write.csv(fundamentals_ds_filter_1, file = "data/fundamentals_ds_filter_3.csv", row.names=FALSE, na="")
```

```{r, column_count_after_nzv_removal, MESSAGE = FALSE}
ncol(fundamentals_ds_filter_1)
nrow(fundamentals_ds_filter_1)
```

### Based on manual analysis of the columns and considering the end goal, remove the variables which logically do not make sense or might have correlattion with existing set of variables.
```{r remove_not_required_variables, MESSAGE = FALSE, warning=FALSE }
fundamentals_ds_filter_1 <- subset(fundamentals_ds_filter_1, select = -c(acctstd,src,acodo,acox,
                                                                          aox,capxv,ceql,cibegni,cicurr,cidergl,
                                                                          cimii,cipen,cisecgl,citotal,cshfd,cshpri,dclo,dcpstk,
                                                                          dltis,dlto,dltr,do,dp,dpvieb,drc,
                                                                          dv,dvc,epsfx,epspx,exre,
                                                                          fopox,ibadj,ibc,ibcom,ibmii,
                                                                          invch,ivaco,ivao,lco,
                                                                          lcox,lcoxdr,lct,loxdr,mibn,
                                                                          mibt,mii,msa,niadj,np,oprepsx,pnca,ppent,ppeveb,
                                                                          recco,rectr,sale,spced,spceeps,
                                                                          tstkc,txbco,txbcof,txdb,
                                                                          txdbca,txdi,
                                                                          txditc,txndb,xopr,exchg,costat,
                                                                          ceoso,cfoso,idbflag,naics,sic,stko))
write.csv(fundamentals_ds_filter_1, file = "data/fundamentals_ds_filter_4.csv", row.names=FALSE, na="")
```

```{r column_count_after_col_removal, MESSAGE = FALSE, warning=FALSE}
colnames(fundamentals_ds_filter_1)
nrow(fundamentals_ds_filter_1)
#fundamentals_ds_filter_1
```


### Filter and create a seperate dataset for restatement 
#### Remove all the columns which contain 10% or greater NA or null values  
```{r filter_data_re_stmt, MESSAGE = FALSE, warning=FALSE}
fundamentals_restmt_ds_filter <- fundamentals_ds %>%
  filter(datafmt == 'SUMM_STD' & gsector == gsector_sjm)
  #filter(datafmt == 'SUMM_STD')
  
std_cols <- colnames(fundamentals_ds_filter_1)
fundamentals_restmt_ds_filter <- subset(fundamentals_restmt_ds_filter, select = c(std_cols))
fundamentals_restmt_ds_filter <- fundamentals_restmt_ds_filter[ lapply( fundamentals_restmt_ds_filter, function(x) sum(is.na(x)) / length(x) ) < 0.1 ]
summary(fundamentals_restmt_ds_filter)
nrow(fundamentals_restmt_ds_filter)
```

### Add restement variables and the magnitude as pecentages 
```{r compare_restatement_rows, MESSAGE = FALSE, warning=FALSE}

sample_restmt_ds_filter <- fundamentals_restmt_ds_filter #%>%
  #filter(gvkey == 1076)
sample_ds_filter <- fundamentals_ds_filter_1 #%>%
  #filter(gvkey == 1076)
#nrow(sample_restmt_ds_filter)
#nrow(sample_ds_filter)
#head(sample_restmt_ds_filter)
#head(sample_ds_filter)

fundamentals_ds_filter_1$restmt_at <- 0
fundamentals_ds_filter_1$restmt_at_mag <- 0.0

fundamentals_ds_filter_1$restmt_capx <- 0
fundamentals_ds_filter_1$restmt_capx_mag <- 0.0

fundamentals_ds_filter_1$restmt_cogs <- 0
fundamentals_ds_filter_1$restmt_cogs_mag <- 0.0

fundamentals_ds_filter_1$restmt_dltt <- 0
fundamentals_ds_filter_1$restmt_dltt_mag <- 0.0

fundamentals_ds_filter_1$restmt_epsfi <- 0
fundamentals_ds_filter_1$restmt_epsfi_mag <- 0.0

fundamentals_ds_filter_1$restmt_epspi <- 0
fundamentals_ds_filter_1$restmt_epspi_mag <- 0.0

fundamentals_ds_filter_1$restmt_ib <- 0
fundamentals_ds_filter_1$restmt_ib_mag <- 0.0

fundamentals_ds_filter_1$restmt_ni <- 0
fundamentals_ds_filter_1$restmt_ni_mag <- 0.0

fundamentals_ds_filter_1$restmt_nopi <- 0
fundamentals_ds_filter_1$restmt_nopi_mag <- 0.0

fundamentals_ds_filter_1$restmt_pi <- 0
fundamentals_ds_filter_1$restmt_pi_mag <- 0.0

fundamentals_ds_filter_1$restmt_reuna <- 0
fundamentals_ds_filter_1$restmt_reuna_mag <- 0.0

fundamentals_ds_filter_1$restmt_seq <- 0
fundamentals_ds_filter_1$restmt_seq_mag <- 0.0

fundamentals_ds_filter_1$restmt_teq <- 0
fundamentals_ds_filter_1$restmt_teq_mag <- 0.0

fundamentals_ds_filter_1$restmt_txt <- 0
fundamentals_ds_filter_1$restmt_txt_mag <- 0.0

fundamentals_ds_filter_1$restmt_wcap <- 0
fundamentals_ds_filter_1$restmt_wcap_mag <- 0.0

#fundamentals_ds_filter_1$restmt_ci <- 0
#fundamentals_ds_filter_1$restmt_ci_mag <- 0.0

fundamentals_ds_filter_1$restmt_xint <- 0
fundamentals_ds_filter_1$restmt_xint_mag <- 0.0

fundamentals_ds_filter_1$restmt_xsga <- 0
fundamentals_ds_filter_1$restmt_xsga_mag <- 0.0

fundamentals_ds_filter_1$restmt_dvpsp_f <- 0
fundamentals_ds_filter_1$restmt_dvpsp_f_mag <- 0.0

fundamentals_ds_filter_1$restmt_dvpsx_f <- 0
fundamentals_ds_filter_1$restmt_dvpsx_f_mag <- 0.0

for (row in 1:nrow(sample_restmt_ds_filter)){
  restmt_item_gvkey  <- as.integer(sample_restmt_ds_filter[row, "gvkey"])
  restmt_item_fyear  <- sample_restmt_ds_filter[row, "fyear"]
  restmt_item_at  <- sample_restmt_ds_filter[row, "at"]
  restmt_item_capx  <- sample_restmt_ds_filter[row, "capx"]
  restmt_item_cogs  <- sample_restmt_ds_filter[row, "cogs"]
  restmt_item_dltt  <- sample_restmt_ds_filter[row, "dltt"]
  restmt_item_epsfi  <- sample_restmt_ds_filter[row, "epsfi"]
  restmt_item_epspi  <- sample_restmt_ds_filter[row, "epspi"]
  
  restmt_item_ib  <- sample_restmt_ds_filter[row, "ib"]
  restmt_item_ni  <- sample_restmt_ds_filter[row, "ni"]
  restmt_item_nopi  <- sample_restmt_ds_filter[row, "nopi"]
  restmt_item_pi  <- sample_restmt_ds_filter[row, "pi"]
  restmt_item_reuna  <- sample_restmt_ds_filter[row, "reuna"]
  restmt_item_seq  <- sample_restmt_ds_filter[row, "seq"]
  restmt_item_teq  <- sample_restmt_ds_filter[row, "teq"]
  restmt_item_txt  <- sample_restmt_ds_filter[row, "txt"]
  restmt_item_wcap  <- sample_restmt_ds_filter[row, "wcap"]
  
  restmt_item_xint  <- sample_restmt_ds_filter[row, "xint"]
  restmt_item_xsga  <- sample_restmt_ds_filter[row, "xsga"]
  restmt_item_dvpsp_f  <- sample_restmt_ds_filter[row, "dvpsp_f"]
  restmt_item_dvpsx_f  <- sample_restmt_ds_filter[row, "dvpsx_f"]

  row_count <- as.integer(nrow(subset(fundamentals_ds_filter_1, gvkey == restmt_item_gvkey & fyear == restmt_item_fyear)))
  
  if (row_count > 0){
    fundamental_stmt_row <- fundamentals_ds_filter_1 %>%
       filter(gvkey == restmt_item_gvkey & fyear == restmt_item_fyear)

    stmt_item_gvkey  <- fundamental_stmt_row["gvkey"]
    stmt_item_fyear  <- fundamental_stmt_row["fyear"]
    stmt_item_at  <- fundamental_stmt_row["at"]
    stmt_item_capx  <- fundamental_stmt_row["capx"]
    stmt_item_cogs  <- fundamental_stmt_row["cogs"]
    stmt_item_dltt  <- fundamental_stmt_row["dltt"]
    stmt_item_epsfi  <- fundamental_stmt_row["epsfi"]
    stmt_item_epspi  <- fundamental_stmt_row["epspi"]
    stmt_item_ib  <- fundamental_stmt_row["ib"]
    stmt_item_ni  <- fundamental_stmt_row["ni"]
    stmt_item_nopi  <- fundamental_stmt_row["nopi"]
    stmt_item_pi  <- fundamental_stmt_row["pi"]
    stmt_item_reuna  <- fundamental_stmt_row["reuna"]
    stmt_item_seq  <- fundamental_stmt_row["seq"]
    stmt_item_teq  <- fundamental_stmt_row["teq"]
    stmt_item_txt  <- fundamental_stmt_row["txt"]
    stmt_item_wcap  <- fundamental_stmt_row["wcap"]
    stmt_item_xint  <- fundamental_stmt_row["xint"]
    stmt_item_xsga  <- fundamental_stmt_row["xsga"]
    stmt_item_dvpsp_f  <- fundamental_stmt_row["dvpsp_f"]
    stmt_item_dvpsx_f  <- fundamental_stmt_row["dvpsx_f"]

    

    if (!is.na(restmt_item_at) & !is.na(stmt_item_at) & stmt_item_at != 0 & restmt_item_at != stmt_item_at){
      fundamentals_ds_filter_1$restmt_at[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_at - stmt_item_at)/stmt_item_at) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_at_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_capx) & !is.na(stmt_item_capx) & restmt_item_capx != stmt_item_capx){
      fundamentals_ds_filter_1$restmt_capx[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_capx == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_capx - stmt_item_capx)/stmt_item_capx) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_capx_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_cogs) & !is.na(stmt_item_cogs) & restmt_item_cogs != stmt_item_cogs){
      fundamentals_ds_filter_1$restmt_cogs[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_cogs == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_cogs - stmt_item_cogs)/stmt_item_cogs) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_cogs_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_dltt) & !is.na(stmt_item_dltt) & restmt_item_dltt != stmt_item_dltt){
      fundamentals_ds_filter_1$restmt_dltt[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_dltt == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_dltt - stmt_item_dltt)/stmt_item_dltt) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_dltt_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_epsfi) & !is.na(stmt_item_epsfi) & restmt_item_epsfi != stmt_item_epsfi){
      fundamentals_ds_filter_1$restmt_epsfi[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_epsfi == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_epsfi - stmt_item_epsfi)/stmt_item_epsfi) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_epsfi_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_epspi) & !is.na(stmt_item_epspi) & restmt_item_epspi != stmt_item_epspi){
      fundamentals_ds_filter_1$restmt_epspi[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_epspi == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_epspi - stmt_item_epspi)/stmt_item_epspi) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_epspi_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }


    if (!is.na(restmt_item_ib) & !is.na(stmt_item_ib) & restmt_item_ib != stmt_item_ib){
      fundamentals_ds_filter_1$restmt_ib[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_ib - stmt_item_ib)/stmt_item_ib) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_ib_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_ni) & !is.na(stmt_item_ni) & restmt_item_ni != stmt_item_ni){
      fundamentals_ds_filter_1$restmt_ni[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_ni - stmt_item_ni)/stmt_item_ni) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_ni_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_nopi) & !is.na(stmt_item_nopi) & restmt_item_nopi != stmt_item_nopi){
      fundamentals_ds_filter_1$restmt_nopi[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_nopi == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_nopi - stmt_item_nopi)/stmt_item_nopi) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_nopi_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_pi) & !is.na(stmt_item_pi) & restmt_item_pi != stmt_item_pi){
      fundamentals_ds_filter_1$restmt_pi[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_pi - stmt_item_pi)/stmt_item_pi) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_pi_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_reuna) & !is.na(stmt_item_reuna) & restmt_item_reuna != stmt_item_reuna){
      fundamentals_ds_filter_1$restmt_reuna[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_reuna - stmt_item_reuna)/stmt_item_reuna) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_reuna_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_seq) & !is.na(stmt_item_seq) & restmt_item_seq != stmt_item_seq){
      fundamentals_ds_filter_1$restmt_seq[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_seq - stmt_item_seq)/stmt_item_seq) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_seq_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_teq) & !is.na(stmt_item_teq) & restmt_item_teq != stmt_item_teq){
      fundamentals_ds_filter_1$restmt_teq[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_teq - stmt_item_teq)/stmt_item_teq) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_teq_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_txt) & !is.na(stmt_item_txt) & restmt_item_txt != stmt_item_txt){
      fundamentals_ds_filter_1$restmt_txt[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      if (stmt_item_txt == 0.0){
        magnitude  <- 100.00
      }
      else{
        magnitude  <- ((restmt_item_txt - stmt_item_txt)/stmt_item_txt) * 100.0
      }
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_txt_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }

    if (!is.na(restmt_item_wcap) & !is.na(stmt_item_wcap) & restmt_item_wcap != stmt_item_wcap){
      fundamentals_ds_filter_1$restmt_wcap[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_wcap - stmt_item_wcap)/stmt_item_wcap) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_wcap_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }
    
    if (!is.na(restmt_item_xint) & !is.na(stmt_item_xint) & stmt_item_xint != 0 & restmt_item_xint != stmt_item_xint){
      fundamentals_ds_filter_1$restmt_xint[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_xint - stmt_item_xint)/stmt_item_xint) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_xint_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }
    
    if (!is.na(restmt_item_xsga) & !is.na(stmt_item_xsga) & restmt_item_xsga != stmt_item_xsga){
      fundamentals_ds_filter_1$restmt_xsga[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_xsga - stmt_item_xsga)/stmt_item_xsga) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_xsga_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }
    
    if (!is.na(restmt_item_dvpsp_f) & !is.na(stmt_item_dvpsp_f) & restmt_item_dvpsp_f != stmt_item_dvpsp_f){
      fundamentals_ds_filter_1$restmt_dvpsp_f[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_dvpsp_f - stmt_item_dvpsp_f)/stmt_item_dvpsp_f) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_dvpsp_f_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }
    
    if (!is.na(restmt_item_dvpsx_f) & !is.na(stmt_item_dvpsx_f) & restmt_item_dvpsx_f != stmt_item_dvpsx_f){
      fundamentals_ds_filter_1$restmt_dvpsx_f[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- 1
      magnitude  <- ((restmt_item_dvpsx_f - stmt_item_dvpsx_f)/stmt_item_dvpsx_f) * 100.0
      magnitude  <- as.double(round(magnitude, digits = 3))
      fundamentals_ds_filter_1$restmt_dvpsx_f_mag[fundamentals_ds_filter_1$gvkey == restmt_item_gvkey & fundamentals_ds_filter_1$fyear == restmt_item_fyear] <- magnitude
    }
    
  }
}
#head(fundamentals_ds_filter_1)
nrow(fundamentals_ds_filter_1)
```

### Removing aco == NA as all those rows do not have any data
```{r summary_ds_with_restamt_flags, MESSAGE = FALSE, warning=FALSE}
nrow(fundamentals_ds_filter_1)
fundamentals_ds_filter_2 <- subset(fundamentals_ds_filter_1, !is.na(aco))
nrow(fundamentals_ds_filter_2)
fundamentals_ds_filter_2[is.na(fundamentals_ds_filter_2)] <- 0
summary(fundamentals_ds_filter_2)
```

```{r get_col_names}
colnames(fundamentals_ds_filter_2)
nrow(fundamentals_ds_filter_2)
```


### The dataset contains company wise data across financial year, this dataset needs to be consolidate to single ror for each of the company 
### So group the dataset by gvkey and summarize all variables 
```{r final_ds_preparation, MESSAGE = FALSE, warning=FALSE}
final_ds_initial <- fundamentals_ds_filter_2 %>%
  group_by(gvkey,tic) %>%
  summarize(
    aco = mean(aco),
    acominc = mean(acominc),
    act = mean(act),
    ao = mean(ao),
    aocidergl = mean(aocidergl),
    aocipen = mean(aocipen),
    aodo = mean(aodo),
    aoloch = mean(aoloch),
    ap = mean(ap),
    aqc = mean(aqc),
    at = mean(at),
    bkvlps = mean(bkvlps),
    caps = mean(caps),
    capx = mean(capx),
    ceq = mean(ceq),
    ceqt = mean(ceqt),
    ch = mean(ch),
    che = mean(che),
    chech = mean(chech),
    ci = mean(ci),
    cogs = mean(cogs),
    cshi = mean(cshi),
    csho = mean(csho),
    cstk = mean(cstk),
    cstkcv = mean(cstkcv),
    dd1 = mean(dd1),
    dilavx = mean(dilavx),
    dlc = mean(dlc),
    dltt = mean(dltt),
    dm = mean(dm),
    dn = mean(dn),
    dpact = mean(dpact),
    dpc = mean(dpc),
    dvt = mean(dvt),
    ebit = mean(ebit),
    ebitda = mean(ebitda),
    epsfi = mean(epsfi),
    epspi = mean(epspi),
    fiao = mean(fiao),
    fincf = mean(fincf),
    fopo = mean(fopo),
    gdwl = mean(gdwl),
    gp = mean(gp),
    ib = mean(ib),
    icapt = mean(icapt),
    intan = mean(intan),
    intano = mean(intano),
    invt = mean(invt),
    ivch = mean(ivch),
    ivncf = mean(ivncf),
    ivst = mean(ivst),
    lo = mean(lo),
    lse = mean(lse),
    lt = mean(lt),
    ni = mean(ni),
    nopi = mean(nopi),
    nopio = mean(nopio),
    oancf = mean(oancf),
    oiadp = mean(oiadp),
    oibdp = mean(oibdp),
    opeps = mean(opeps),
    pi = mean(pi),
    ppegt = mean(ppegt),
    re = mean(re),
    reajo = mean(reajo),
    rect = mean(rect),
    recta = mean(recta),
    reuna = mean(reuna),
    revt = mean(revt),
    seq	 = 	mean(	seq	),
    siv	 = 	mean(	siv	),
    spce = mean(spce),
    spi = mean(spi),
    sppiv = mean(sppiv),
    sstk = mean(sstk),
    teq = mean(teq),
    tstk = mean(tstk),
    tstkn = mean(tstkn),
    txp = mean(txp),
    txr = mean(txr),
    txt = mean(txt),
    wcap = mean(wcap),
    xint = mean(xint),
    restmt_at = mean(restmt_at),
    restmt_at_mag = mean(restmt_at_mag),
    restmt_capx = mean(restmt_capx),
    restmt_capx_mag = mean(restmt_capx_mag),
    restmt_cogs = mean(restmt_cogs),
    restmt_cogs_mag = mean(restmt_cogs_mag),
    restmt_dltt = mean(restmt_dltt),
    restmt_dltt_mag = mean(restmt_dltt_mag),
    restmt_epsfi = mean(restmt_epsfi),
    restmt_epsfi_mag = mean(restmt_epsfi_mag),
    restmt_epspi = mean(restmt_epspi),
    restmt_epspi_mag = mean(restmt_epspi_mag),
    restmt_ib = mean(restmt_ib),
    restmt_ib_mag = mean(restmt_ib_mag),
    restmt_ni = mean(restmt_ni),
    restmt_ni_mag = mean(restmt_ni_mag),
    restmt_nopi = mean(restmt_nopi),
    restmt_nopi_mag = mean(restmt_nopi_mag),
    restmt_pi = mean(restmt_pi),
    restmt_pi_mag = mean(restmt_pi_mag),
    restmt_reuna = mean(restmt_reuna),
    restmt_reuna_mag = mean(restmt_reuna_mag),
    restmt_seq = mean(restmt_seq),
    restmt_seq_mag = mean(restmt_seq_mag),
    restmt_teq = mean(restmt_teq),
    restmt_teq_mag = mean(restmt_teq_mag),
    restmt_txt = mean(restmt_txt),
    restmt_txt_mag = mean(restmt_txt_mag),
    restmt_wcap = mean(restmt_wcap),
    restmt_wcap_mag = mean(restmt_wcap_mag),
    
    restmt_xint = mean(restmt_xint),
    restmt_xint_mag = mean(restmt_xint_mag),
    
    restmt_xsga = mean(restmt_xsga),
    restmt_xsga_mag = mean(restmt_xsga_mag),
    
    restmt_dvpsp_f = mean(restmt_dvpsp_f),
    restmt_dvpsp_f_mag = mean(restmt_dvpsp_f_mag),
    
    restmt_dvpsx_f = mean(restmt_dvpsx_f),
    restmt_dvpsx_f_mag = mean(restmt_dvpsx_f_mag)
    
  )

summary(final_ds_initial)
nrow(final_ds_initial)
```

### Adjust the restement values, in case if the restate percentage is greater than 50% then mark restatement as 1 or else mark as 0 For 0/false restatement magnitude is marked as 0.
```{r adjust_restatement_flag, MESSAGE = FALSE, warning=FALSE}
final_ds_initial_1 <- final_ds_initial 

for (row in 1:nrow(final_ds_initial_1)){
  row_item_gvkey  <- as.integer(final_ds_initial_1[row, "gvkey"])
  
  restmt_at <- final_ds_initial_1[row, "restmt_at"]
  restmt_at_mag <- final_ds_initial_1[row, "restmt_at_mag"]
  if (restmt_at >= 0.5){
    restmt_at <- 1
    restmt_at_mag <- as.double(restmt_at_mag)
  }
  else{
    restmt_at <- 0
    restmt_at_mag <- 0.0
  }
  final_ds_initial_1$restmt_at[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_at
  final_ds_initial_1$restmt_at_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_at_mag
  
  restmt_capx <- final_ds_initial_1[row, "restmt_capx"]
  restmt_capx_mag <- final_ds_initial_1[row, "restmt_capx_mag"]
  if (restmt_capx >= 0.5){
    restmt_capx <- 1
    restmt_capx_mag <- as.double(restmt_capx_mag)
  }
  else{
    restmt_capx <- 0
    restmt_capx_mag <- 0.0
  }
  final_ds_initial_1$restmt_capx[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_capx
  final_ds_initial_1$restmt_capx_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_capx_mag
  
  restmt_cogs <- final_ds_initial_1[row, "restmt_cogs"]
  restmt_cogs_mag <- final_ds_initial_1[row, "restmt_cogs_mag"]
  if (restmt_cogs >= 0.5){
    restmt_cogs <- 1
    restmt_cogs_mag <- as.double(restmt_cogs_mag)
  }
  else{
    restmt_cogs <- 0
    restmt_cogs_mag <- 0.0
  }
  final_ds_initial_1$restmt_cogs[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_cogs
  final_ds_initial_1$restmt_cogs_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_cogs_mag
  
  restmt_dltt <- final_ds_initial_1[row, "restmt_dltt"]
  restmt_dltt_mag <- final_ds_initial_1[row, "restmt_dltt_mag"]
  if (restmt_dltt >= 0.5){
    restmt_dltt <- 1
    restmt_dltt_mag <- as.double(restmt_dltt_mag)
  }
  else{
    restmt_dltt <- 0
    restmt_dltt_mag <- 0.0
  }
  final_ds_initial_1$restmt_dltt[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_dltt
  final_ds_initial_1$restmt_dltt_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_dltt_mag
  
  
  restmt_epsfi <- final_ds_initial_1[row, "restmt_epsfi"]
  restmt_epsfi_mag <- final_ds_initial_1[row, "restmt_epsfi_mag"]
  if (restmt_epsfi >= 0.5){
    restmt_epsfi <- 1
    restmt_epsfi_mag <- as.double(restmt_epsfi_mag)
  }
  else{
    restmt_epsfi <- 0
    restmt_epsfi_mag <- 0.0
  }
  final_ds_initial_1$restmt_epsfi[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_epsfi
  final_ds_initial_1$restmt_epsfi_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_epsfi_mag
  
  
  restmt_epspi <- final_ds_initial_1[row, "restmt_epspi"]
  restmt_epspi_mag <- final_ds_initial_1[row, "restmt_epspi_mag"]
  if (restmt_epspi >= 0.5){
    restmt_epspi <- 1
    restmt_epspi_mag <- as.double(restmt_epspi_mag)
  }
  else{
    restmt_epspi <- 0
    restmt_epspi_mag <- 0.0
  }
  final_ds_initial_1$restmt_epspi[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_epspi
  final_ds_initial_1$restmt_epspi_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_epspi_mag
  
  restmt_ib <- final_ds_initial_1[row, "restmt_ib"]
  restmt_ib_mag <- final_ds_initial_1[row, "restmt_ib_mag"]
  if (restmt_ib >= 0.5){
    restmt_ib <- 1
    restmt_ib_mag <- as.double(restmt_ib_mag)
  }
  else{
    restmt_ib <- 0
    restmt_ib_mag <- 0.0
  }
  final_ds_initial_1$restmt_ib[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_ib
  final_ds_initial_1$restmt_ib_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_ib_mag
  
  restmt_ni <- final_ds_initial_1[row, "restmt_ni"]
  restmt_ni_mag <- final_ds_initial_1[row, "restmt_ni_mag"]
  if (restmt_ni >= 0.5){
    restmt_ni <- 1
    restmt_ni_mag <- as.double(restmt_ni_mag)
  }
  else{
    restmt_ni <- 0
    restmt_ni_mag <- 0.0
  }
  final_ds_initial_1$restmt_ni[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_ni
  final_ds_initial_1$restmt_ni_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_ni_mag
  
  restmt_nopi <- final_ds_initial_1[row, "restmt_nopi"]
  restmt_nopi_mag <- final_ds_initial_1[row, "restmt_nopi_mag"]
  if (restmt_nopi >= 0.5){
    restmt_nopi <- 1
    restmt_nopi_mag <- as.double(restmt_nopi_mag)
  }
  else{
    restmt_nopi <- 0
    restmt_nopi_mag <- 0.0
  }
  final_ds_initial_1$restmt_nopi[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_nopi
  final_ds_initial_1$restmt_nopi_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_nopi_mag
  
  restmt_pi <- final_ds_initial_1[row, "restmt_pi"]
  restmt_pi_mag <- final_ds_initial_1[row, "restmt_pi_mag"]
  if (restmt_pi >= 0.5){
    restmt_pi <- 1
    restmt_pi_mag <- as.double(restmt_pi_mag)
  }
  else{
    restmt_pi <- 0
    restmt_pi_mag <- 0.0
  }
  final_ds_initial_1$restmt_pi[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_pi
  final_ds_initial_1$restmt_pi_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_pi_mag
  
  restmt_reuna <- final_ds_initial_1[row, "restmt_reuna"]
  restmt_reuna_mag <- final_ds_initial_1[row, "restmt_reuna_mag"]
  if (restmt_reuna >= 0.5){
    restmt_reuna <- 1
    restmt_reuna_mag <- as.double(restmt_reuna_mag)
  }
  else{
    restmt_reuna <- 0
    restmt_reuna_mag <- 0.0
  }
  final_ds_initial_1$restmt_reuna[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_reuna
  final_ds_initial_1$restmt_reuna_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_reuna_mag
  
  restmt_seq <- final_ds_initial_1[row, "restmt_seq"]
  restmt_seq_mag <- final_ds_initial_1[row, "restmt_seq_mag"]
  if (restmt_seq >= 0.5){
    restmt_seq <- 1
    restmt_seq_mag <- as.double(restmt_seq_mag)
  }
  else{
    restmt_seq <- 0
    restmt_seq_mag <- 0.0
  }
  final_ds_initial_1$restmt_seq[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_seq
  final_ds_initial_1$restmt_seq_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_seq_mag
  
  restmt_teq <- final_ds_initial_1[row, "restmt_teq"]
  restmt_teq_mag <- final_ds_initial_1[row, "restmt_teq_mag"]
  if (restmt_teq >= 0.5){
    restmt_teq <- 1
    restmt_teq_mag <- as.double(restmt_teq_mag)
  }
  else{
    restmt_teq <- 0
    restmt_teq_mag <- 0.0
  }
  final_ds_initial_1$restmt_teq[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_teq
  final_ds_initial_1$restmt_teq_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_teq_mag
  
  restmt_txt <- final_ds_initial_1[row, "restmt_txt"]
  restmt_txt_mag <- final_ds_initial_1[row, "restmt_txt_mag"]
  if (restmt_txt >= 0.5){
    restmt_txt <- 1
    restmt_txt_mag <- as.double(restmt_txt_mag)
  }
  else{
    restmt_txt <- 0
    restmt_txt_mag <- 0.0
  }
  final_ds_initial_1$restmt_txt[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_txt
  final_ds_initial_1$restmt_txt_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_txt_mag
  
  restmt_wcap <- final_ds_initial_1[row, "restmt_wcap"]
  restmt_wcap_mag <- final_ds_initial_1[row, "restmt_wcap_mag"]
  if (restmt_wcap >= 0.5){
    restmt_wcap <- 1
    restmt_wcap_mag <- as.double(restmt_wcap_mag)
  }
  else{
    restmt_wcap <- 0
    restmt_wcap_mag <- 0.0
  }
  final_ds_initial_1$restmt_wcap[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_wcap
  final_ds_initial_1$restmt_wcap_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_wcap_mag
  
  restmt_xint <- final_ds_initial_1[row, "restmt_xint"]
  restmt_xint_mag <- final_ds_initial_1[row, "restmt_xint_mag"]
  if (restmt_xint >= 0.5){
    restmt_xint <- 1
    restmt_xint_mag <- as.double(restmt_xint_mag)
  }
  else{
    restmt_xint <- 0
    restmt_xint_mag <- 0.0
  }
  final_ds_initial_1$restmt_xint[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_xint
  final_ds_initial_1$restmt_xint_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_xint_mag
  
  restmt_xsga <- final_ds_initial_1[row, "restmt_xsga"]
  restmt_xsga_mag <- final_ds_initial_1[row, "restmt_xsga_mag"]
  if (restmt_xsga >= 0.5){
    restmt_xsga <- 1
    restmt_xsga_mag <- as.double(restmt_xsga_mag)
  }
  else{
    restmt_xsga <- 0
    restmt_xsga_mag <- 0.0
  }
  final_ds_initial_1$restmt_xsga[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_xsga
  final_ds_initial_1$restmt_xsga_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_xsga_mag
  
  restmt_dvpsp_f <- final_ds_initial_1[row, "restmt_dvpsp_f"]
  restmt_dvpsp_f_mag <- final_ds_initial_1[row, "restmt_dvpsp_f_mag"]
  if (restmt_dvpsp_f >= 0.5){
    restmt_dvpsp_f <- 1
    restmt_dvpsp_f_mag <- as.double(restmt_dvpsp_f_mag)
  }
  else{
    restmt_dvpsp_f <- 0
    restmt_dvpsp_f_mag <- 0.0
  }
  final_ds_initial_1$restmt_dvpsp_f[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_dvpsp_f
  final_ds_initial_1$restmt_dvpsp_f_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_dvpsp_f_mag
  
  restmt_dvpsx_f <- final_ds_initial_1[row, "restmt_dvpsx_f"]
  restmt_dvpsx_f_mag <- final_ds_initial_1[row, "restmt_dvpsx_f_mag"]
  if (restmt_dvpsx_f >= 0.5){
    restmt_dvpsx_f <- 1
    restmt_dvpsx_f_mag <- as.double(restmt_dvpsx_f_mag)
  }
  else{
    restmt_dvpsx_f <- 0
    restmt_dvpsx_f_mag <- 0.0
  }
  final_ds_initial_1$restmt_dvpsx_f[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_dvpsx_f
  final_ds_initial_1$restmt_dvpsx_f_mag[final_ds_initial_1$gvkey == row_item_gvkey] <- restmt_dvpsx_f_mag
}

restmt_var_ds <- subset(final_ds_initial_1, select = c(gvkey,
                                                     restmt_at,restmt_at_mag,
                                                      restmt_capx,restmt_capx_mag,
                                                      restmt_cogs, restmt_cogs_mag,
                                                      restmt_dltt, restmt_dltt_mag,
                                                      restmt_epsfi, restmt_epsfi_mag,
                                                      restmt_epspi, restmt_epspi_mag,
                                                      restmt_ib, restmt_ib_mag,
                                                      restmt_ni, restmt_ni_mag,
                                                      restmt_nopi, restmt_nopi_mag,
                                                      restmt_pi, restmt_pi_mag,
                                                      restmt_reuna, restmt_reuna_mag,
                                                      restmt_seq, restmt_seq_mag,
                                                      restmt_teq, restmt_teq_mag,
                                                      restmt_txt, restmt_txt_mag,
                                                      restmt_wcap, restmt_wcap_mag,
                                                     
                                                     restmt_xint, restmt_xint_mag,
                                                     restmt_xsga, restmt_xsga_mag,
                                                     restmt_dvpsp_f, restmt_dvpsp_f_mag,
                                                     restmt_dvpsx_f, restmt_dvpsx_f_mag
                                                     ))

summary(restmt_var_ds)
```

```{r final_dataset, MESSAGE = FALSE, warning=FALSE}
final_ds_initial_2 <- final_ds_initial_1
summary(final_ds_initial_2)
nrow(final_ds_initial_2)
write.csv(final_ds_initial_2, file = "data/final_ds_initial_2.csv", row.names=FALSE, na="")
```

### This first step where all the correlated variables are idenfied and then removed. This will reduce the Collinearity. 
```{r identify_correlation, MESSAGE = FALSE, warning=FALSE}
cor_matrix_ds <- subset(final_ds_initial_2, select = -c(gvkey,tic, aodo,seq,ivch,nopio,spce,reuna,dilavx,ebitda,cshi,epsfi, 
                                              ib,pi,
                                              oiadp,oibdp,gdwl))
cor_matrix <- cor(cor_matrix_ds)
cor_matrix %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  group_by(value) %>%
  filter(row_number()==1)
#corrplot(cor_matrix, method = "ellipse")
#ncol(cor_matrix_ds)
```

```{r fundamentals_final_ds, MESSAGE = FALSE, warning=FALSE}
fundamentals_final_ds <- subset(final_ds_initial_2, select = -c(aodo,seq,ivch,nopio,spce,reuna,dilavx,ebitda,cshi,epsfi, 
                                              ib,pi,
                                              oiadp,oibdp,gdwl))
summary(fundamentals_final_ds)
nrow(fundamentals_final_ds)
```

### Loan stocks data file 
```{r load_stocks_data, MESSAGE = FALSE, warning=FALSE}
stocks_init_ds <- read.csv("./data/Stocks_DS.csv", na.strings=c(""," "))
nrow(stocks_init_ds)
```

### From Stocks daat file choose only certain variables 
### prccd	==> Price - Close - Daily
### prchd	==> Price - High - Daily
### prcld	==> Price - Low - Daily
### prcod	==> Price - Open - Daily
### trfd	==> Daily Total Return Factor
### cshtrd ==> Trading Volume - Daily
```{r stocks_data_filter, warning=FALSE}
names(stocks_init_ds)[names(stocks_init_ds) == "誰..gvkey"] <- "gvkey"
stocks_init_limited_cols <- subset(stocks_init_ds, select = c(gvkey,cshtrd,prccd,prchd,prcld,prcod,trfd))
stocks_init_limited_cols <- stocks_init_limited_cols[!is.na(stocks_init_limited_cols$cshtrd)&!is.na(stocks_init_limited_cols$prccd)
                                                     &!is.na(stocks_init_limited_cols$prchd)&!is.na(stocks_init_limited_cols$prcld)
                                                     &!is.na(stocks_init_limited_cols$trfd),]
stocks_init_limited_cols$prcod[is.na(stocks_init_limited_cols$prcod)] <- (stocks_init_limited_cols$prchd + stocks_init_limited_cols$prcld)/2

stocks_grouped_data <- stocks_init_limited_cols %>%
   group_by(gvkey) %>%
   summarize(
     cshtrd_m = mean(cshtrd),
     prccd_m = mean(prccd),
     prchd_m = mean(prchd),
     prcld_m = mean(prcld),
     prcod_m = mean(prcod),
     trfd_m = mean(trfd)
   )
ncol(stocks_init_ds)
```


### Merge the stocks dataset with final dataset.
```{r, merge_fundamental_stocks_data, warning=FALSE}

#fundamental_stocks_data <- fundamentals_final_ds %>%
#  inner_join(stocks_grouped_data, by = 'gvkey')
#summary(fundamental_stocks_data)


for (row in 1:nrow(fundamentals_final_ds)){
  row_item_gvkey  <- as.integer(fundamentals_final_ds[row, "gvkey"])
  
  specific_stock <- stocks_grouped_data %>%
    filter(gvkey == row_item_gvkey) 
  
  if (nrow(specific_stock) > 0){
    specific_stock <- head(specific_stock, 1)
    fundamentals_final_ds$cshtrd_m[final_ds_initial_1$gvkey == row_item_gvkey] <- specific_stock$cshtrd_m
    fundamentals_final_ds$prccd_m[final_ds_initial_1$gvkey == row_item_gvkey] <- specific_stock$prccd_m
    fundamentals_final_ds$prchd_m[final_ds_initial_1$gvkey == row_item_gvkey] <- specific_stock$prchd_m
    fundamentals_final_ds$prcld_m[final_ds_initial_1$gvkey == row_item_gvkey] <- specific_stock$prcld_m
    fundamentals_final_ds$prcod_m[final_ds_initial_1$gvkey == row_item_gvkey] <- specific_stock$prcod_m
    fundamentals_final_ds$trfd_m[final_ds_initial_1$gvkey == row_item_gvkey] <- specific_stock$trfd_m
  }
}
summary(fundamentals_final_ds)
nrow(fundamentals_final_ds)

fundamental_stocks_data <- fundamentals_final_ds[!is.na(fundamentals_final_ds$cshtrd_m) &!is.na(fundamentals_final_ds$prccd_m)
                                                     &!is.na(fundamentals_final_ds$prchd_m) &!is.na(fundamentals_final_ds$prcld_m)
                                                     &!is.na(fundamentals_final_ds$prcod_m) &!is.na(fundamentals_final_ds$trfd_m),]
```

```{r identify_correlation_1, MESSAGE = FALSE, warning=FALSE}
# cor_matrix_ds <- subset(fundamental_stocks_data, select = -c(gvkey,tic))
# cor_matrix <- cor(cor_matrix_ds)
# cor_matrix %>%
#   as.data.frame() %>%
#   mutate(var1 = rownames(.)) %>%
#   gather(var2, value, -var1) %>%
#   arrange(desc(value)) %>%
#   group_by(value) %>%
#   filter(row_number()==1)
```

```{r drop_cols_based_on_above_correlations, warning=FALSE}
fundamental_stocks_data_final <- fundamental_stocks_data
#summary(fundamental_stocks_data_final)
nrow(fundamental_stocks_data_final)
```



### Load securities dataset
```{r load_securities_data, MESSAGE = FALSE, warning=FALSE}
securities_init_ds <- read.csv("./data/Securities_DS.csv", na.strings=c(""," "))
names(securities_init_ds)[names(securities_init_ds) == "誰..gvkey"] <- "gvkey"
ncol(securities_init_ds)
securities_init_ds_1 <- subset(securities_init_ds, select = -c(iid,isalrt,primiss,ajexm,
                                                               spgim,spiim,spmim,cheqvm,curcddvm,dvpsxm,
                                                               sphcusip,sphiid, sphmid,sphname,sphsec,sphtic,sphvg,sph100,
                                                               cyear,mkvalincl,exchg,tpci,city,
                                                               conml,costat,ggroup,gind, gsubind,loc,naics,sic,state, curcdm, 
                                                               navm,adrrm,rawpm,rawxm,cshoq,csfsm,
                                                               datadate,tic,conm,cmth
                                                               ))


#summary(securities_init_ds_1)
```

### Choose following variables from the securities dataset 
### trfm	  ==> Monthly Total Return Factor
### dvrate	==> Dividend Rate - Monthly
```{r merge_stock_data, warning=FALSE, message=FALSE}
fund_stock_securities_ds <- fundamental_stocks_data_final 
securities_init_ds_2 <- securities_init_ds_1 %>%
  filter(!is.na(trfm) & !is.na(trt1m)) %>%
  group_by(gvkey) %>%
  summarise(
    trfm_m = mean(trfm)
    )

fund_stock_securities_ds$trfm_m <- NA
for (row in 1:nrow(fund_stock_securities_ds)){
  row_item_gvkey  <- as.integer(fund_stock_securities_ds[row, "gvkey"])
  specific_security <- securities_init_ds_2 %>%
    filter(gvkey == row_item_gvkey) 
  if (nrow(specific_security) > 0){
    security_row <- head(specific_security, 1)
    trfm_m <- as.numeric(security_row$trfm_m)  
    fund_stock_securities_ds$trfm_m[fund_stock_securities_ds$gvkey == row_item_gvkey] <- trfm_m
  }
}



securities_init_ds_3 <- securities_init_ds_1 %>%
  filter(!is.na(dvrate)) %>%
  group_by(gvkey) %>%
  summarise(
    dvrate_m = mean(dvrate)
    )

fund_stock_securities_ds$dvrate_m <- NA
for (row in 1:nrow(fund_stock_securities_ds)){
  row_item_gvkey  <- as.integer(fund_stock_securities_ds[row, "gvkey"])
  specific_security <- securities_init_ds_3 %>%
    filter(gvkey == row_item_gvkey) 
  if (nrow(specific_security) > 0){
    security_row <- head(specific_security, 1)
    dvrate_m <- as.numeric(security_row$dvrate_m)  
    fund_stock_securities_ds$dvrate_m[fund_stock_securities_ds$gvkey == row_item_gvkey] <- dvrate_m
  }
}


#summary(fund_stock_securities_ds)
```

### Load ratings dataset.
### Load this variable splticrm	==> S&P Domestic Long Term Issuer Credit Rating
### splticrm	- This is categorical variable with unique ratings, each of the rating are given numeric values.
### Highest rating gets high numeric values and as rating decreases the numerica value assigned decreases. 
```{r load_ratings_data, MESSAGE = FALSE, warning=FALSE}
ratings_init_ds <- read.csv("./data/Ratings_DS.csv", na.strings=c("", " "))
names(ratings_init_ds)[names(ratings_init_ds) == "誰..gvkey"] <- "gvkey"
ratings_init_ds$datadate <- as.Date(ratings_init_ds$datadate, "%m/%d/%Y")
ratings_init_ds$splticrm = factor(ratings_init_ds$splticrm, levels=c(levels(ratings_init_ds$splticrm), "NR"))
ratings_init_ds$splticrm[is.na(ratings_init_ds$splticrm)] = "NR"

ratings_init_ds$splticrm_num_value <- 0
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "AAA"] <-  100
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "AA"] <-   90
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "AA-"] <-  85
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "A+"] <-   80
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "A"] <-    75
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "A-"] <-   70


ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "BBB+"] <-   65
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "BBB"] <-    60
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "BBB-"] <-   55
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "BB+"] <-    50
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "BB"] <-   45
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "BB-"] <-   40

ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "B+"] <-   35
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "B"] <-   30
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "B-"] <-   25


ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "CCC+"] <-   20
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "CCC"] <-   19
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "CCC-"] <-   18
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "CC"] <-   17

ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "D"] <-   10
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "SD"] <-   10
ratings_init_ds$splticrm_num_value[ratings_init_ds$splticrm == "NR"] <-   0
ratings_init_ds$splticrm_num_value <- factor(ratings_init_ds$splticrm_num_value)
#levels(ratings_init_ds$splticrm)
#str(ratings_init_ds)
```

### Merge the ratings dataset and fundamentals dataset, by assigning the rating from ratings dataset to each company using gvkey  
### Along with rating, additional variable is added which indicates whether the rating has 
###     1. Increased
###     2. Decreased 
###     3. NOCHANGE
```{r, consolidate_ds_fundamental_stock_rating, warning=FALSE, message=FALSE}
fund_stock_securities_rating_ds <- fund_stock_securities_ds #%>%
  #filter(gvkey == 1078)
fund_stock_securities_rating_ds$sp_rating <- "NOTRATED" 
rated_companies <- ratings_init_ds %>%
  filter(splticrm != "NR")

for (row in 1:nrow(fund_stock_securities_rating_ds)){
  row_item_gvkey  <- as.integer(fund_stock_securities_rating_ds[row, "gvkey"])

  specific_rating <- rated_companies %>%
    filter(gvkey == row_item_gvkey) %>%
    arrange(datadate)
  if (nrow(specific_rating) > 0){
    first_row <- head(specific_rating, 1)
    last_row <- tail(specific_rating, 1)
    start_value <- as.integer(first_row$splticrm_num_value)
    end_value <- as.integer(last_row$splticrm_num_value)
    if (start_value == end_value){
      fund_stock_securities_rating_ds$sp_rating[fund_stock_securities_rating_ds$gvkey == row_item_gvkey] <- "NoCHANGE"
    }else if (start_value < end_value){
      fund_stock_securities_rating_ds$sp_rating[fund_stock_securities_rating_ds$gvkey == row_item_gvkey] <- "INCREASED"
    }else if (start_value > end_value){
      fund_stock_securities_rating_ds$sp_rating[fund_stock_securities_rating_ds$gvkey == row_item_gvkey] <- "DECREASED"
    }
  }
}
fund_stock_securities_rating_ds$sp_rating <- factor(fund_stock_securities_rating_ds$sp_rating)
#summary(fund_stock_securities_rating_ds)
```

### Load Sca filing dataset for the target variable 
```{r load_sca_filings_data, MESSAGE = FALSE, warning=FALSE}
sca_fillings_ds <- read.csv("./data/SCA_Filings_and_Settlements.csv", na.strings=c(""," "))
sca_fillings_ds$SettlementAmount = gsub("\\$", "", sca_fillings_ds$SettlementAmount)
sca_fillings_ds$SettlementAmount = as.numeric(gsub("\\,", "", sca_fillings_ds$SettlementAmount))
#summary(sca_fillings_ds)
ncol(sca_fillings_ds)
```

### Add the target variable litigated to the dataset under analysis.
### From each of the conany record identify if has any sca filing from the sca fileing dataset,if entry exists then mark for that company lititgated = true or else if the record ### does not exist in sca filing  for that particular comapny then mark litigate attribute as false.
### Also in case of litigation indentify if there is any settlement amount and add same to main dataset. In case if multiple settlement amount exists for particular company then ### take the maximum settlement amount.  
```{r merge_sca_filings_data, MESSAGE = FALSE, warning=FALSE}
fund_stock_securities_rating_ds$litigated <- 0
fund_stock_securities_rating_ds$litigation_settlement <- NA
fund_stock_securities_rating_ds_1 <- fund_stock_securities_rating_ds 

for (row in 1:nrow(fund_stock_securities_rating_ds_1)){
  row_item_tic  <- lapply(fund_stock_securities_rating_ds_1[row, "tic"], as.character)
  row_item_gvkey  <- as.integer(fund_stock_securities_rating_ds[row, "gvkey"])
  specific_sca_filings <- sca_fillings_ds %>%
    filter(Ticker == row_item_tic) 
  if (nrow(specific_sca_filings) > 0){
    fund_stock_securities_rating_ds$litigated[fund_stock_securities_rating_ds$gvkey == row_item_gvkey] <- 1
    specific_sca_filings_max <- specific_sca_filings %>%
    filter(!is.na(SettlementAmount)) %>%
    arrange(SettlementAmount)
    
   if (nrow(specific_sca_filings_max) > 0){
      last_row <- tail(specific_sca_filings_max, 1)
      settlement_amount <- as.numeric(last_row$SettlementAmount)
      fund_stock_securities_rating_ds$litigation_settlement[fund_stock_securities_rating_ds$gvkey == row_item_gvkey] <- settlement_amount
   }
  }
}
fund_stock_securities_rating_ds$litigated <- as.factor(fund_stock_securities_rating_ds$litigated)
#summary(fund_stock_securities_rating_ds)
```

```{r final_ds_column_names, MESSAGE = FALSE, warning=FALSE}
#colnames(fund_stock_securities_rating_ds)
```


### Based on manual review. idetify the columns that need to be removed from the final dataset.
```{r clean_up_not_required_columns, MESSAGE = FALSE, warning=FALSE}
fund_stock_securities_rating_ds_final <- subset(fund_stock_securities_rating_ds, select = -c(aocidergl, aocipen, ceqt, cstkcv, dd1,
                                                    dpc,icapt, intan, intano, ivncf, ivst,
                                                    lo, lse, opeps, reajo, recta,
                                                    spi, tstkn, txp, txr,
                                                    restmt_epsfi_mag, restmt_epsfi,
                                                    restmt_pi, restmt_pi_mag,
                                                    restmt_seq, restmt_seq_mag,
                                                    restmt_xsga, restmt_xsga_mag,
                                                    restmt_dvpsp_f, restmt_dvpsp_f_mag, 
                                                    restmt_dvpsx_f, restmt_dvpsx_f_mag
                                                    ))
#summary(fund_stock_securities_rating_ds_final)
```

```{r identify_correlation_2, MESSAGE = FALSE, warning=FALSE}
# cor_matrix_ds <- subset(fund_stock_securities_rating_ds_final, select = -c(gvkey,tic, sp_rating, litigated))
# cor_matrix <- cor(cor_matrix_ds)
# cor_matrix %>%
#   as.data.frame() %>%
#   mutate(var1 = rownames(.)) %>%
#   gather(var2, value, -var1) %>%
#   arrange(desc(value)) %>%
#   group_by(value) %>%
#   filter(row_number() == 1)
```

# DE ==> Debt to equity ratio
# wc ==> Working capital ratio
# pe ==> Pricing to earning ratio
# ROE ==> Return on Equity
```{r add_financial_ratios, MESSAGE = FALSE, warning=FALSE}
fund_stock_securities_rating_ds_final$epspi[fund_stock_securities_rating_ds_final$epspi == 0] <- 0.000001
fund_stock_securities_rating_ds_final$lt[fund_stock_securities_rating_ds_final$lt == 0] <- 0.000001
fund_stock_securities_rating_ds_final$teq[fund_stock_securities_rating_ds_final$teq == 0] <- 0.000001
fund_stock_securities_rating_ds_final$pe_ratio <- fund_stock_securities_rating_ds_final$prccd_m/fund_stock_securities_rating_ds_final$epspi
fund_stock_securities_rating_ds_final$wc_ratio <- fund_stock_securities_rating_ds_final$act/fund_stock_securities_rating_ds_final$lt
fund_stock_securities_rating_ds_final$de_ratio <- fund_stock_securities_rating_ds_final$lt/fund_stock_securities_rating_ds_final$teq
fund_stock_securities_rating_ds_final$roe_ratio <- fund_stock_securities_rating_ds_final$ni/fund_stock_securities_rating_ds_final$teq
#temp <- subset(fund_stock_securities_rating_ds_final, select = c(gvkey,tic, epspi, prccd_m,pe_ratio, act, lt,wc_ratio, teq,de_ratio, ni, roe_ratio))
#head(temp)
trfm_median <- median(as.numeric(fund_stock_securities_rating_ds_final$trfm_m),na.rm=TRUE)
fund_stock_securities_rating_ds_final$trfm_m[is.na(fund_stock_securities_rating_ds_final$trfm_m)] <- trfm_median
```

### From the final dataset identify correlated variables 
```{r identify_correlation_3, MESSAGE = FALSE, warning=FALSE}
cor_matrix_ds <- subset(fund_stock_securities_rating_ds_final, select = -c(gvkey,tic, sp_rating, litigated))
cor_matrix <- cor(cor_matrix_ds)
cor_matrix %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  group_by(value) %>%
  filter(row_number() == 1)

```

#### Remove Highly correlated variables, in general if correlation value is > 0.9 and check the correlation matrix values
```{r identify_correlation_4, MESSAGE = FALSE, warning=FALSE}
cor_matrix_ds <- subset(fund_stock_securities_rating_ds_final, select = -c(gvkey,tic, sp_rating, litigated, prchd_m, prcld_m, ni, restmt_ib_mag, ceq, oancf, cogs, ppegt,
                                                                           lt, ci, restmt_dltt_mag, invt, che, ap, at, xint, gp, act, txt, capx,
                                                                           dm, dn, dpact,fiao,fincf,sppiv,fopo)) #latest removed variables
cor_matrix <- cor(cor_matrix_ds)
cor_matrix %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  group_by(value) %>%
  filter(row_number() == 1)

```



### Remove Highly correlated variables from above analysis.
```{r ds_final_rows_summary_analysis, MESSAGE = FALSE, warning=FALSE}

fund_stock_securities_rating_ds_final_10 <- subset(fund_stock_securities_rating_ds_final, select = -c(prchd_m, prcld_m, ni, restmt_ib_mag, 
                                                                                                      ceq, oancf, cogs, ppegt,
                                                                                                      lt, ci, restmt_dltt_mag, invt, che, 
                                                                                                      ap, at, xint, gp, act, txt, capx,
                                                                                                      dm, dn, dpact,fiao,fincf,sppiv,fopo))  #latest removed variables
#nrow(fund_stock_securities_rating_ds_final)
#summary(fund_stock_securities_rating_ds_final)
```

### Convert all categorical variables to factor
```{r convert_categorical_to_factor, MESSAGE = FALSE, warning=FALSE}
ds_final <- fund_stock_securities_rating_ds_final_10
ds_final <- subset(ds_final, select = -c(dvrate_m)) #, litigation_settlement))
ds_final$restmt_at <- as.factor(ds_final$restmt_at)
ds_final$restmt_capx <- as.factor(ds_final$restmt_capx)
ds_final$restmt_cogs <- as.factor(ds_final$restmt_cogs)
ds_final$restmt_dltt <- as.factor(ds_final$restmt_dltt)
ds_final$restmt_epspi <- as.factor(ds_final$restmt_epspi)
ds_final$restmt_ib <- as.factor(ds_final$restmt_ib)
ds_final$restmt_ni <- as.factor(ds_final$restmt_ni)
ds_final$restmt_nopi <- as.factor(ds_final$restmt_nopi)
ds_final$restmt_reuna <- as.factor(ds_final$restmt_reuna)
ds_final$restmt_teq <- as.factor(ds_final$restmt_teq)
ds_final$restmt_txt <- as.factor(ds_final$restmt_txt)
ds_final$restmt_wcap <- as.factor(ds_final$restmt_wcap)
ds_final$restmt_xint <- as.factor(ds_final$restmt_xint)
ds_final$sp_rating <- as.factor(ds_final$sp_rating)
ds_final$litigated <- as.factor(ds_final$litigated)
```



### Perform target ecoding to all the categorical variables 
```{r encode_categorical_variables, MESSAGE = FALSE, warning=FALSE}
# split the data into training and (held-out) test sets
training_ind <- createDataPartition(ds_final$litigated,
p = 0.75,
list = FALSE,
times = 1)
training_set <- ds_final[training_ind, ]
test_set <- ds_final[-training_ind, ]

nrow(training_set)
nrow(test_set)
threshold <- 250
#head(training_set$litigated)

threshold <- 250
target_enc_train <- function(variable, level) {
  training_set$litigated <- as.numeric(as.vector(training_set$litigated))
  train_avg_target <- colMeans(training_set[, "litigated"])
  if (nrow(training_set[training_set[, variable]==level, ])==0) {
    return(train_avg_target)
  } else {
    level_num_obs <- nrow(training_set[training_set[, variable]==level,])
    level_avg_target <- colMeans(training_set[training_set[, variable]==level, "litigated"])
    return((level_num_obs*level_avg_target+threshold*train_avg_target)/(level_num_obs+threshold))
  }
}

sp_rating_target <- mapply(target_enc_train, variable = "sp_rating", level = levels(training_set$sp_rating), USE.NAMES = FALSE)
names(sp_rating_target) <- levels(training_set$sp_rating)
training_set$sp_rating_target <- 0
for (level in levels(training_set$sp_rating)) {
  training_set[training_set[, "sp_rating"]==level, "sp_rating_target"] <- sp_rating_target[level]
}

test_set$sp_rating_target <- 0
for (level in levels(training_set$sp_rating)) {
  test_set[test_set[, "sp_rating"]==level, "sp_rating_target"] <- sp_rating_target[level]
}


restmt_at_target <- mapply(target_enc_train, variable = "restmt_at", level = levels(training_set$restmt_at), USE.NAMES = FALSE)
names(restmt_at_target) <- levels(training_set$restmt_at)
training_set$restmt_at_target <- 0
for (level in levels(training_set$restmt_at)) {
  training_set[training_set[, "restmt_at"]==level, "restmt_at_target"] <- restmt_at_target[level]
}

test_set$restmt_at_target <- 0
for (level in levels(training_set$restmt_at)) {
  test_set[test_set[, "restmt_at"]==level, "restmt_at_target"] <- restmt_at_target[level]
}

restmt_capx_target <- mapply(target_enc_train, variable = "restmt_capx", level = levels(training_set$restmt_capx), USE.NAMES = FALSE)
names(restmt_capx_target) <- levels(training_set$restmt_capx)
training_set$restmt_capx_target <- 0
for (level in levels(training_set$restmt_capx)) {
  training_set[training_set[, "restmt_capx"]==level, "restmt_capx_target"] <- restmt_capx_target[level]
}

test_set$restmt_capx_target <- 0
for (level in levels(training_set$restmt_capx)) {
  test_set[test_set[, "restmt_capx"]==level, "restmt_capx_target"] <- restmt_capx_target[level]
}

restmt_cogs_target <- mapply(target_enc_train, variable = "restmt_cogs", level = levels(training_set$restmt_cogs), USE.NAMES = FALSE)
names(restmt_cogs_target) <- levels(training_set$restmt_cogs)
training_set$restmt_cogs_target <- 0
for (level in levels(training_set$restmt_cogs)) {
  training_set[training_set[, "restmt_cogs"]==level, "restmt_cogs_target"] <- restmt_cogs_target[level]
}

test_set$restmt_cogs_target <- 0
for (level in levels(training_set$restmt_cogs)) {
  test_set[test_set[, "restmt_cogs"]==level, "restmt_cogs_target"] <- restmt_cogs_target[level]
}

restmt_dltt_target <- mapply(target_enc_train, variable = "restmt_dltt", level = levels(training_set$restmt_dltt), USE.NAMES = FALSE)
names(restmt_dltt_target) <- levels(training_set$restmt_dltt)
training_set$restmt_dltt_target <- 0
for (level in levels(training_set$restmt_dltt)) {
  training_set[training_set[, "restmt_dltt"]==level, "restmt_dltt_target"] <- restmt_dltt_target[level]
}

test_set$restmt_dltt_target <- 0
for (level in levels(training_set$restmt_dltt)) {
  test_set[test_set[, "restmt_dltt"]==level, "restmt_dltt_target"] <- restmt_dltt_target[level]
}

restmt_epspi_target <- mapply(target_enc_train, variable = "restmt_epspi", level = levels(training_set$restmt_epspi), USE.NAMES = FALSE)
names(restmt_epspi_target) <- levels(training_set$restmt_epspi)
training_set$restmt_epspi_target <- 0
for (level in levels(training_set$restmt_epspi)) {
  training_set[training_set[, "restmt_epspi"]==level, "restmt_epspi_target"] <- restmt_epspi_target[level]
}

test_set$restmt_epspi_target <- 0
for (level in levels(training_set$restmt_epspi)) {
  test_set[test_set[, "restmt_epspi"]==level, "restmt_epspi_target"] <- restmt_epspi_target[level]
}

restmt_ib_target <- mapply(target_enc_train, variable = "restmt_ib", level = levels(training_set$restmt_ib), USE.NAMES = FALSE)
names(restmt_ib_target) <- levels(training_set$restmt_ib)
training_set$restmt_ib_target <- 0
for (level in levels(training_set$restmt_ib)) {
  training_set[training_set[, "restmt_ib"]==level, "restmt_ib_target"] <- restmt_ib_target[level]
}

test_set$restmt_ib_target <- 0
for (level in levels(training_set$restmt_ib)) {
  test_set[test_set[, "restmt_ib"]==level, "restmt_ib_target"] <- restmt_ib_target[level]
}


restmt_ni_target <- mapply(target_enc_train, variable = "restmt_ni", level = levels(training_set$restmt_ni), USE.NAMES = FALSE)
names(restmt_ni_target) <- levels(training_set$restmt_ni)
training_set$restmt_ni_target <- 0
for (level in levels(training_set$restmt_ni)) {
  training_set[training_set[, "restmt_ni"]==level, "restmt_ni_target"] <- restmt_ni_target[level]
}

test_set$restmt_ni_target <- 0
for (level in levels(training_set$restmt_ni)) {
  test_set[test_set[, "restmt_ni"]==level, "restmt_ni_target"] <- restmt_ni_target[level]
}

restmt_nopi_target <- mapply(target_enc_train, variable = "restmt_nopi", level = levels(training_set$restmt_nopi), USE.NAMES = FALSE)
names(restmt_nopi_target) <- levels(training_set$restmt_nopi)
training_set$restmt_nopi_target <- 0
for (level in levels(training_set$restmt_nopi)) {
  training_set[training_set[, "restmt_nopi"]==level, "restmt_nopi_target"] <- restmt_nopi_target[level]
}

test_set$restmt_nopi_target <- 0
for (level in levels(training_set$restmt_nopi)) {
  test_set[test_set[, "restmt_nopi"]==level, "restmt_nopi_target"] <- restmt_nopi_target[level]
}


restmt_reuna_target <- mapply(target_enc_train, variable = "restmt_reuna", level = levels(training_set$restmt_reuna), USE.NAMES = FALSE)
names(restmt_reuna_target) <- levels(training_set$restmt_reuna)
training_set$restmt_reuna_target <- 0
for (level in levels(training_set$restmt_reuna)) {
  training_set[training_set[, "restmt_reuna"]==level, "restmt_reuna_target"] <- restmt_reuna_target[level]
}

test_set$restmt_reuna_target <- 0
for (level in levels(training_set$restmt_reuna)) {
  test_set[test_set[, "restmt_reuna"]==level, "restmt_reuna_target"] <- restmt_reuna_target[level]
}

restmt_teq_target <- mapply(target_enc_train, variable = "restmt_teq", level = levels(training_set$restmt_teq), USE.NAMES = FALSE)
names(restmt_teq_target) <- levels(training_set$restmt_teq)
training_set$restmt_teq_target <- 0
for (level in levels(training_set$restmt_teq)) {
  training_set[training_set[, "restmt_teq"]==level, "restmt_teq_target"] <- restmt_teq_target[level]
}

test_set$restmt_teq_target <- 0
for (level in levels(training_set$restmt_teq)) {
  test_set[test_set[, "restmt_teq"]==level, "restmt_teq_target"] <- restmt_teq_target[level]
}

restmt_txt_target <- mapply(target_enc_train, variable = "restmt_txt", level = levels(training_set$restmt_txt), USE.NAMES = FALSE)
names(restmt_txt_target) <- levels(training_set$restmt_txt)
training_set$restmt_txt_target <- 0
for (level in levels(training_set$restmt_txt)) {
  training_set[training_set[, "restmt_txt"]==level, "restmt_txt_target"] <- restmt_txt_target[level]
}

test_set$restmt_txt_target <- 0
for (level in levels(training_set$restmt_txt)) {
  test_set[test_set[, "restmt_txt"]==level, "restmt_txt_target"] <- restmt_txt_target[level]
}

restmt_wcap_target <- mapply(target_enc_train, variable = "restmt_wcap", level = levels(training_set$restmt_wcap), USE.NAMES = FALSE)
names(restmt_wcap_target) <- levels(training_set$restmt_wcap)
training_set$restmt_wcap_target <- 0
for (level in levels(training_set$restmt_wcap)) {
  training_set[training_set[, "restmt_wcap"]==level, "restmt_wcap_target"] <- restmt_wcap_target[level]
}

test_set$restmt_wcap_target <- 0
for (level in levels(training_set$restmt_wcap)) {
  test_set[test_set[, "restmt_wcap"]==level, "restmt_wcap_target"] <- restmt_wcap_target[level]
}

restmt_xint_target <- mapply(target_enc_train, variable = "restmt_xint", level = levels(training_set$restmt_xint), USE.NAMES = FALSE)
names(restmt_xint_target) <- levels(training_set$restmt_xint)
training_set$restmt_xint_target <- 0
for (level in levels(training_set$restmt_xint)) {
  training_set[training_set[, "restmt_xint"]==level, "restmt_xint_target"] <- restmt_xint_target[level]
}

test_set$restmt_xint_target <- 0
for (level in levels(training_set$restmt_xint)) {
  test_set[test_set[, "restmt_xint"]==level, "restmt_xint_target"] <- restmt_xint_target[level]
}
```

```{r save_target_company}
nrow(training_set)
nrow(test_set)
target_company_row <- training_set %>% 
  filter(gvkey == target_company_gv_key)

if (nrow(target_company_row) > 0) {
  training_set <- subset(training_set, gvkey != target_company_gv_key)
} else {
  
  target_company_row <- test_set %>% 
    filter(gvkey == target_company_gv_key)
  test_set <- subset(test_set, gvkey != target_company_gv_key)
}

nrow(training_set)
nrow(test_set)
target_company_row
```

### Remove the restatement variables which have near zero variance
```{r keep_encoded_required_variables, MESSAGE = FALSE, warning=FALSE}
training_subset_ds_final_1 <- subset(training_set, select = -c(sp_rating,
                                                              restmt_at,
                                                              restmt_capx,
                                                              restmt_cogs,
                                                              restmt_dltt,
                                                              restmt_epspi,
                                                              restmt_ib,
                                                              restmt_ni,
                                                              restmt_nopi,
                                                              restmt_reuna,
                                                              restmt_teq,
                                                              restmt_txt,
                                                              restmt_wcap,
                                                              restmt_xint))


test_subset_ds_final_1 <- subset(test_set, select = -c(sp_rating,
                                                              restmt_at,
                                                              restmt_capx,
                                                              restmt_cogs,
                                                              restmt_dltt,
                                                              restmt_epspi,
                                                              restmt_ib,
                                                              restmt_ni,
                                                              restmt_nopi,
                                                              restmt_reuna,
                                                              restmt_teq,
                                                              restmt_txt,
                                                              restmt_wcap,
                                                              restmt_xint))

target_company_row_1 <- subset(target_company_row, select = -c(sp_rating,
                                                              restmt_at,
                                                              restmt_capx,
                                                              restmt_cogs,
                                                              restmt_dltt,
                                                              restmt_epspi,
                                                              restmt_ib,
                                                              restmt_ni,
                                                              restmt_nopi,
                                                              restmt_reuna,
                                                              restmt_teq,
                                                              restmt_txt,
                                                              restmt_wcap,
                                                              restmt_xint))


training_subset_ds_final_1 <- training_subset_ds_final_1 %>%
  relocate(litigated, .after = last_col())

test_subset_ds_final_1 <- test_subset_ds_final_1 %>%
  relocate(litigated, .after = last_col())

target_company_row_final <- target_company_row_1 %>%
  relocate(litigated, .after = last_col())


training_subset_ds_final_1 <- training_subset_ds_final_1 %>%
  relocate(litigation_settlement , .after = last_col())

test_subset_ds_final_1 <- test_subset_ds_final_1 %>%
  relocate(litigation_settlement, .after = last_col())

target_company_row_final <- target_company_row_final %>%
  relocate(litigation_settlement, .after = last_col())

training_subset_ds_final <- training_subset_ds_final_1
test_subset_ds_final <- test_subset_ds_final_1

ncol(training_subset_ds_final)
summary(training_subset_ds_final)
target_company_row_final
training_subset_ds_final_settlement <- training_subset_ds_final
test_subset_ds_final_settlement <- test_subset_ds_final
```



### Scale the variables. 
```{r, SCALE_VARIABLES, MESSAGE = FALSE, warning=FALSE}
num_var_start_index <- 3
num_var_end_index <- ncol(training_subset_ds_final) - 2
target_var_index <- ncol(training_subset_ds_final) - 1

num_var_start_index
num_var_end_index
target_var_index
test_subset_ds_final[, num_var_start_index:num_var_end_index] <- scale(test_subset_ds_final[, num_var_start_index:num_var_end_index], 
                                       center = apply(training_subset_ds_final[, num_var_start_index:num_var_end_index], 2, mean), 
                                       scale = apply(training_subset_ds_final[, num_var_start_index:num_var_end_index], 2, sd))

#target_company_row_final
target_company_row_final[, num_var_start_index:num_var_end_index] = scale(target_company_row_final[, num_var_start_index:num_var_end_index], 
                                       center = apply(training_subset_ds_final[, num_var_start_index:num_var_end_index], 2, mean), 
                                       scale = apply(training_subset_ds_final[, num_var_start_index:num_var_end_index], 2, sd))

training_subset_ds_final[, num_var_start_index:num_var_end_index] <- scale(training_subset_ds_final[, num_var_start_index:num_var_end_index])

levels(training_subset_ds_final$litigated)[levels(training_subset_ds_final$litigated) == 1] <- "Yes"
levels(training_subset_ds_final$litigated)[levels(training_subset_ds_final$litigated) == 0] <- "No"

target_company_row_final
summary(training_subset_ds_final)
```

```{r over_sampling_data_class_imbalance_1}
training_subset_ds_final_orig <- training_subset_ds_final
```

```{r list_all_columns}
colnames(training_subset_ds_final)
```

```{r over_sampling_data_class_imbalance}
training_subset_ds_final <- training_subset_ds_final_orig
table(training_subset_ds_final$litigated)
training_subset_ds_final_no <- training_subset_ds_final %>%
  filter(litigated == 'No')
no_count <- nrow(training_subset_ds_final_no)
training_subset_ds_final_yes <- training_subset_ds_final %>%
  filter(litigated == 'Yes')
yes_count <- nrow(training_subset_ds_final_yes)
#training_subset_ds_final %>%
#  arrange(gvkey)

table(training_subset_ds_final$litigated)

# Sampling both (Over and under combination)
training_subset_ds_final <- ovun.sample(litigated ~ 
                                                    ., 
                                        data = training_subset_ds_final, method = "both", N = no_count + yes_count, p = 0.35, seed = 222, na.action = na.pass)$data

##### Sampling both (Over)
#training_subset_ds_final <- ovun.sample(litigated ~., data = training_subset_ds_final, method = "over", N = no_count*2)$data

##### Sampling both (Under)
#training_subset_ds_final  <- ovun.sample(litigated ~., data = training_subset_ds_final, method = "under", N = yes_count*2)$data

##### Manual sample adjustment
# litigated_no <- training_subset_ds_final %>% 
#   filter(litigated == 'No')
# 
# litigated_yes <- training_subset_ds_final %>% 
#   filter(litigated == 'Yes')
# 
# litigated_no_sample_index <- sample(nrow(litigated_no), 97)
# litigated_no_sample <- litigated_no[litigated_no_sample_index, ]
# #litigated_no_sample
# 
# training_subset_ds_final <- bind_rows(litigated_yes, litigated_no_sample) 
# 
# rows <- sample(nrow(training_subset_ds_final))
# training_subset_ds_final <- training_subset_ds_final[rows, ]

table(training_subset_ds_final$litigated)
#summary(training_subset_ds_final)
training_subset_ds_final
```


```{r  logisitic_regression_glm_1, MESSAGE = FALSE, warning=FALSE}
glm_control <- trainControl(
  method = "cv",
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE 
)
set.seed(123)
glm_model_1 <- train(litigated ~
                      aco + acominc + ao + aoloch + aqc + bkvlps + caps + 
                      ch + chech + csho + cstk + dlc + dltt + dvt +
                      ebit + epspi + nopi + re + rect + revt + siv +
                      sstk + teq + tstk + wcap +
                      cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                      wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                      restmt_at_target + 
                      restmt_capx_target + 
                      restmt_dltt_target + 
                      restmt_wcap_target + 
                      restmt_teq_target + restmt_teq_mag +
                      restmt_cogs_target + restmt_cogs_mag +
                      restmt_nopi_target + restmt_nopi_mag
                      , 
                      data = training_subset_ds_final, method = "glm", family = "binomial", trControl = glm_control)
class_probabilities <- predict(glm_model_1, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
glm_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
glm_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
glm_confusion_matrix <- confusionMatrix(factor(glm_class_probabilities_litigated), factor(glm_litigated), positive = "Yes")
glm_confusion_matrix
glm_accuracy_1 <- glm_confusion_matrix$overall["Accuracy"]
glm_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
glm_rocr_roc <- performance(glm_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(glm_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
glm_rocr_auc <- performance(glm_rocr_pred, measure = "auc")
glm_auc_1 <- glm_rocr_auc@y.values[[1]]
glm_auc_1
summary(glm_model_1)
```


```{r  logisitic_regression_glm_2, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
glm_model_2 <- train(litigated ~
                      acominc + 
                      bkvlps + 
                      caps + 
                      dlc + 
                      dvt + 
                      nopi + 
                      re + 
                      rect + 
                      revt + 
                      tstk + 
                      wcap + 
                      prccd_m + 
                      prcod_m + 
                      trfm_m  + 
                      pe_ratio + 
                      wc_ratio + 
                      sp_rating_target + 
                      restmt_at_target + 
                      restmt_capx_target + 
                      restmt_dltt_target + 
                      restmt_cogs_target + 
                      restmt_nopi_target + 
                      restmt_nopi_mag
                      , 
                      data = training_subset_ds_final, method = "glm", family = "binomial", trControl = glm_control)
class_probabilities <- predict(glm_model_2, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
glm_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
glm_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
glm_confusion_matrix <- confusionMatrix(factor(glm_class_probabilities_litigated), factor(glm_litigated), positive = "Yes")
glm_confusion_matrix
glm_accuracy_2 <- glm_confusion_matrix$overall["Accuracy"]
glm_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
glm_rocr_roc <- performance(glm_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(glm_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
glm_rocr_auc <- performance(glm_rocr_pred, measure = "auc")
glm_auc_2 <- glm_rocr_auc@y.values[[1]]
glm_auc_2
summary(glm_model_2)
```

```{r  logisitic_regression_glm_3, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
glm_model_3 <- train(litigated ~
                      dvt + 
                      nopi + 
                      revt + 
                      tstk + 
                      prccd_m + 
                      trfm_m +  
                      pe_ratio + 
                      wc_ratio + 
                      sp_rating_target
                      , 
                      data = training_subset_ds_final, method = "glm", family = "binomial", trControl = glm_control)
class_probabilities <- predict(glm_model_3, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
glm_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
glm_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
glm_confusion_matrix <- confusionMatrix(factor(glm_class_probabilities_litigated), factor(glm_litigated), positive = "Yes")
glm_confusion_matrix
glm_accuracy_3 <- glm_confusion_matrix$overall["Accuracy"]
glm_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
glm_rocr_roc <- performance(glm_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(glm_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
glm_rocr_auc <- performance(glm_rocr_pred, measure = "auc")
glm_auc_3 <- glm_rocr_auc@y.values[[1]]
glm_auc_3
summary(glm_model_3)
```
```{r  glmnet_1, MESSAGE = FALSE, warning=FALSE}
glmnet_control <- trainControl(
  method = "cv",
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE 
)
set.seed(123)
glmnet_model_1 <- train(litigated ~ 
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                      data = training_subset_ds_final, method = "glmnet", family = "binomial", trControl = glmnet_control)
class_probabilities <- predict(glmnet_model_1, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
glmnet_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
glmnet_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
glmnet_confusion_matrix <- confusionMatrix(factor(glmnet_class_probabilities_litigated), factor(glmnet_litigated), positive = "Yes")
glmnet_confusion_matrix
glmnet_accuracy_1 <- glmnet_confusion_matrix$overall["Accuracy"]
```

```{r  glmnet_2, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
glmnet_model_2 <- train(litigated ~ 
                        aco + acominc + ao + aoloch + aqc + bkvlps + caps + 
                        ch + chech + csho + cstk + dlc + dltt + dvt +
                        ebit + epspi + nopi + re + rect + revt + siv +
                        sstk + teq + tstk + wcap +
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_target + 
                        restmt_capx_target + 
                        restmt_dltt_target + 
                        restmt_wcap_target + 
                        restmt_teq_target + restmt_teq_mag +
                        restmt_cogs_target + restmt_cogs_mag +
                        restmt_nopi_target + restmt_nopi_mag, 
                      data = training_subset_ds_final, method = "glmnet", family = "binomial", trControl = glmnet_control)
class_probabilities <- predict(glmnet_model_2, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
glmnet_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
glmnet_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
glmnet_confusion_matrix <- confusionMatrix(factor(glmnet_class_probabilities_litigated), factor(glmnet_litigated), positive = "Yes")
glmnet_confusion_matrix
glmnet_accuracy_2 <- glmnet_confusion_matrix$overall["Accuracy"]
```


```{r  glmnet_3, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
glmnet_model_3 <- train(litigated ~ 
                        bkvlps + 
                        dlc + 
                        dvt + 
                        nopi + 
                        rect + 
                        revt + 
                        tstk + 
                        wcap + 
                        prccd_m +  
                        prcod_m + 
                        pe_ratio + 
                        wc_ratio + 
                        sp_rating_target + 
                        restmt_at_target + 
                        restmt_capx_target + 
                        restmt_nopi_target, 
                      data = training_subset_ds_final, method = "glmnet", family = "binomial", trControl = glmnet_control)
class_probabilities <- predict(glmnet_model_3, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
glmnet_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
glmnet_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
glmnet_confusion_matrix <- confusionMatrix(factor(glmnet_class_probabilities_litigated), factor(glmnet_litigated), positive = "Yes")
glmnet_confusion_matrix
glmnet_accuracy_3 <- glmnet_confusion_matrix$overall["Accuracy"]
```

```{r logistic_regression_glmnet_3_roc_curve, MESSAGE = FALSE, warning=FALSE}
glmnet_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
glmnet_rocr_roc <- performance(glmnet_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(glmnet_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
```
```{r logistic_regression_glmnet_3_auc_r, MESSAGE = FALSE, warning=FALSE}
glmnet_rocr_auc <- performance(glmnet_rocr_pred, measure = "auc")
glmnet_auc <- glmnet_rocr_auc@y.values[[1]]
glmnet_auc
```


```{r gbm_1, MESSAGE = FALSE, warning=FALSE}
gbm_control <- trainControl(method = "repeatedcv", number = 10, repeats = 10)
set.seed(123)
gbm_model_1 <- train(litigated~
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                  data=training_subset_ds_final, 
                  method = "gbm",
                  trControl = gbm_control,
                  verbose = FALSE)

class_probabilities <- predict(gbm_model_1, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes
gbm_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
gbm_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
gbm_confusion_matrix <- confusionMatrix(factor(gbm_class_probabilities_litigated), factor(gbm_litigated), positive = "Yes")
gbm_confusion_matrix
gbm_accuracy_1 <- gbm_confusion_matrix$overall["Accuracy"]
gbm_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
gbm_rocr_roc <- performance(gbm_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(gbm_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
gbm_rocr_auc <- performance(gbm_rocr_pred, measure = "auc")
gbm_auc_1 <- gbm_rocr_auc@y.values[[1]]
gbm_auc_1
summary(gbm_model_1)
```

```{r tree_1, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
tree_model_1 <- rpart(litigated ~ 
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                      data = training_subset_ds_final, method = "class")
class_probabilities <- predict(tree_model_1, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- 1 - class_probabilities[1: nrow(test_subset_ds_final)]
tree_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)

tree_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
tree_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
tree_confusion_matrix <- confusionMatrix(factor(tree_class_probabilities_litigated), factor(tree_litigated), positive = "Yes")
tree_confusion_matrix
tree_accuracy_1 <- tree_confusion_matrix$overall["Accuracy"]

tree_rocr_roc <- performance(tree_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(tree_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)

tree_rocr_auc <- performance(tree_rocr_pred, measure = "auc")
tree_auc <- tree_rocr_auc@y.values[[1]]
tree_auc

```

```{r plot_tree}
rpart.plot(tree_model_1)
```



```{r  logisitic_regression_lda_1, MESSAGE = FALSE, warning=FALSE}
lda_control <- trainControl(
  method = "cv",
  number = 10
)
set.seed(123)
lda_model_1 <- train(litigated ~
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                      data = training_subset_ds_final, method = "lda", family = "binomial", trControl = lda_control)
class_probabilities <- predict(lda_model_1, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_onehot <- class_probabilities$Yes
lda_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
lda_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
lda_confusion_matrix <- confusionMatrix(factor(lda_class_probabilities_litigated), factor(lda_litigated), positive = "Yes")
lda_confusion_matrix
lda_accuracy_1 <- lda_confusion_matrix$overall["Accuracy"]
lda_accuracy_1

lda_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_onehot, test_subset_ds_final$litigated)
lda_rocr_roc <- performance(lda_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(lda_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)

lda_rocr_auc <- performance(lda_rocr_pred, measure = "auc")
lda_auc_1 <- lda_rocr_auc@y.values[[1]]
lda_auc_1
```



```{r random_forest, MESSAGE = FALSE, warning=FALSE}

rf_x <- training_subset_ds_final[,num_var_start_index:num_var_end_index]
rf_y <- training_subset_ds_final[,target_var_index]
recommended_mtry <- sqrt(ncol(rf_x))
rfGrid <- expand.grid(mtry=recommended_mtry)
```

```{r random_forest_1, MESSAGE = FALSE, warning=FALSE}
rfControl <- trainControl(method='repeatedcv', number=10, repeats=3)
set.seed(123)
rf_model_1 <- train(litigated~
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                      data=training_subset_ds_final, 
                      method='rf', 
                      metric='Accuracy', 
                      tuneGrid=rfGrid, 
                      trControl=rfControl, 
                      ntree = 100
                      )

class_probabilities <- predict(rf_model_1, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes

rf_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
rf_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
rf_confusion_matrix <- confusionMatrix(factor(rf_class_probabilities_litigated), factor(rf_litigated), positive = "Yes")
rf_confusion_matrix
rf_accuracy_1 <- rf_confusion_matrix$overall["Accuracy"]
rf_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
rf_rocr_roc <- performance(rf_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(rf_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
rf_rocr_auc <- performance(rf_rocr_pred, measure = "auc")
rf_auc_1 <- rf_rocr_auc@y.values[[1]]
rf_auc_1
calibration_curve <- calibration(litigated ~ class_probabilities_litigated,
                                 data = test_subset_ds_final,
                                 class = 1)
plot(calibration_curve)
plot(rf_model_1$finalModel)
```


```{r random_forest_2, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
rf_model_2 <- train(litigated~
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                      data=training_subset_ds_final, 
                      method='rf', 
                      metric='Accuracy', 
                      tuneGrid=rfGrid, 
                      trControl=rfControl, 
                      ntree = 300
                      )

class_probabilities <- predict(rf_model_2, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes

rf_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
rf_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
rf_confusion_matrix <- confusionMatrix(factor(rf_class_probabilities_litigated), factor(rf_litigated), positive = "Yes")
rf_confusion_matrix
rf_accuracy_2 <- rf_confusion_matrix$overall["Accuracy"]
rf_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
rf_rocr_roc <- performance(rf_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(rf_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
rf_rocr_auc <- performance(rf_rocr_pred, measure = "auc")
rf_auc_2 <- rf_rocr_auc@y.values[[1]]
rf_auc_2
calibration_curve <- calibration(litigated ~ class_probabilities_litigated,
                                 data = test_subset_ds_final,
                                 class = 1)
plot(calibration_curve)
plot(rf_model_2$finalModel)
```

```{r random_forest_3, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
rf_model_3 <- train(litigated~
                        aco + acominc + ao + aoloch + aqc +
                        bkvlps + caps + ch + chech + csho + cstk + dlc +
                        dltt + dvt + ebit + epspi + nopi + re + rect +
                        revt + siv + sstk + teq + tstk + wcap + 
                        cshtrd_m + prccd_m + prcod_m + trfd_m + trfm_m + pe_ratio +
                        wc_ratio + de_ratio + roe_ratio + sp_rating_target +
                        restmt_at_mag + restmt_capx_mag + restmt_cogs_mag + 
                        restmt_epspi_mag + restmt_ni_mag + restmt_nopi_mag + 
                        restmt_reuna_mag + restmt_teq_mag + restmt_txt_mag + 
                        restmt_wcap_mag + restmt_xint_mag + 
                        restmt_at_target + restmt_capx_target + restmt_cogs_target + 
                        restmt_dltt_target + restmt_epspi_target + restmt_ib_target + 
                        restmt_ni_target + restmt_nopi_target + restmt_reuna_target + 
                        restmt_teq_target + restmt_txt_target + restmt_wcap_target + 
                        restmt_xint_target, 
                      data=training_subset_ds_final, 
                      method='rf', 
                      metric='Accuracy', 
                      tuneGrid=rfGrid, 
                      trControl=rfControl, 
                      ntree = 500
                      )

class_probabilities <- predict(rf_model_3, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes

rf_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
rf_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
rf_confusion_matrix <- confusionMatrix(factor(rf_class_probabilities_litigated), factor(rf_litigated), positive = "Yes")
rf_confusion_matrix
rf_accuracy_3 <- rf_confusion_matrix$overall["Accuracy"]
rf_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
rf_rocr_roc <- performance(rf_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(rf_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
rf_rocr_auc <- performance(rf_rocr_pred, measure = "auc")
rf_auc_3 <- rf_rocr_auc@y.values[[1]]
rf_auc_3
calibration_curve <- calibration(litigated ~ class_probabilities_litigated,
                                 data = test_subset_ds_final,
                                 class = 1)
plot(calibration_curve)
plot(rf_model_3$finalModel)
```

```{r feature_importance_random_forest, MESSAGE = FALSE, warning=FALSE}
rf_varImp <- varImp(rf_model_3, type = 2)
plot(rf_varImp, top = 30)
```
```{r list_important_features_random_forest}
var_imp_ds <- rf_varImp$importance
var_imp_ds <- var_imp_ds %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  arrange(desc(Overall))
var_imp_ds
write.csv(var_imp_ds, file = "data/var_imp_ds.csv", row.names=FALSE)
```

```{r random_forest_4, MESSAGE = FALSE, warning=FALSE}
set.seed(123)
rf_model_4 <- train(litigated ~
                      caps + 
                      cshtrd_m + 
                      re + 
                      wcap + 
                      prccd_m + 
                      sstk + 
                      prcod_m + 
                      restmt_nopi_mag + 
                      revt + 
                      aco + 
                      aoloch + 
                      chech + 
                      bkvlps + 
                      ch + 
                      epspi + 
                      rect + 
                      ebit + 
                      cstk + 
                      csho + 
                      teq + 
                      ao + 
                      acominc + 
                      roe_ratio + 
                      dlc + 
                      pe_ratio + 
                      nopi + 
                      trfd_m + 
                      wc_ratio + 
                      de_ratio + 
                      dltt + 
                      tstk + 
                      siv + 
                      aqc + 
                      trfm_m + 
                      dvt + 
                      restmt_epspi_mag + 
                      restmt_cogs_mag
                      , 
                      data=training_subset_ds_final, 
                      method='rf', 
                      metric='Accuracy', 
                      tuneGrid=rfGrid, 
                      trControl=rfControl, 
                      ntree = 500)

class_probabilities <- predict(rf_model_4, newdata = test_subset_ds_final[, -1*c(target_var_index:target_var_index)], type = "prob")
test_subset_ds_final$class_probabilities_litigated <- class_probabilities$Yes

rf_litigated <- ifelse(test_subset_ds_final$litigated == 1, "Yes", "No")
rf_class_probabilities_litigated <- ifelse(test_subset_ds_final$class_probabilities_litigated > 0.50, "Yes", "No")
rf_confusion_matrix <- confusionMatrix(factor(rf_class_probabilities_litigated), factor(rf_litigated), positive = "Yes")
rf_confusion_matrix
rf_accuracy_4 <- rf_confusion_matrix$overall["Accuracy"]


rf_rocr_pred <- prediction(test_subset_ds_final$class_probabilities_litigated, test_subset_ds_final$litigated)
rf_rocr_roc <- performance(rf_rocr_pred, measure = "tpr", x.measure = "fpr")
plot(rf_rocr_roc,
colorize = TRUE,
print.cutoffs.at = seq(0, 1, by = 0.1),
text.adj = c(-0.5, 1),
lwd = 2)
abline(a = 0, b = 1)
rf_rocr_auc <- performance(rf_rocr_pred, measure = "auc")
rf_auc_4 <- rf_rocr_auc@y.values[[1]]
rf_auc_4
calibration_curve <- calibration(litigated ~ class_probabilities_litigated,
                                 data = test_subset_ds_final,
                                 class = 1)
plot(calibration_curve)
plot(rf_model_2$finalModel)
```

```{r, build_data_frame_auc}
auc_df <- data.frame("algorithm_method" = c("glm_1"), "auc" = c(glm_auc_1) * 100, "accuracy" = c(glm_accuracy_1) * 100,  "notes" = c("All variables"))

auc_df_temp <- data.frame("algorithm_method" = c("glm_2"), "auc" = c(glm_auc_2)  * 100, "accuracy" = c(glm_accuracy_2) * 100,  
                          "notes" = c("First Set of Signigficant variables variables"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("glm_3"), "auc" = c(glm_auc_3)  * 100, "accuracy" = c(glm_accuracy_3) * 100,
                          "notes" = c("Second set of signigficant variables variables"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("glmnet_1"), "auc" = c(glmnet_auc)  * 100, "accuracy" = c(glmnet_accuracy_1) * 100,"notes" = c("Default"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("glmnet_1"), "auc" = c(glmnet_auc)  * 100, "accuracy" = c(glmnet_accuracy_2) * 100,"notes" = c("Default"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("glmnet_1"), "auc" = c(glmnet_auc)  * 100, "accuracy" = c(glmnet_accuracy_3) * 100,"notes" = c("Default"))
auc_df <- rbind(auc_df,auc_df_temp)
 
auc_df_temp <- data.frame("algorithm_method" = c("lda_1"), "auc" = c(lda_auc_1) * 100, "accuracy" = c(lda_accuracy_1) * 100,"notes" = c("Default"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("gbm_1"), "auc" = c(gbm_auc_1) * 100, "accuracy" = c(gbm_accuracy_1) * 100,"notes" = c("Default"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("tree_1"), "auc" = c(tree_auc) * 100, "accuracy" = c(tree_accuracy_1) * 100,"notes" = c("Default"))
auc_df <- rbind(auc_df,auc_df_temp)
 
auc_df_temp <- data.frame("algorithm_method" = c("rf_1"), "auc" = c(rf_auc_1) * 100, "accuracy" = c(rf_accuracy_1) * 100,"notes" = c("Trees = 100"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("rf_2"), "auc" = c(rf_auc_2) * 100, "accuracy" = c(rf_accuracy_2) * 100,"notes" = c("Trees = 300"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("rf_3"), "auc" = c(rf_auc_3) * 100, "accuracy" = c(rf_accuracy_3) * 100,"notes" = c("Trees = 500"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df_temp <- data.frame("algorithm_method" = c("rf_4"), "auc" = c(rf_auc_4) * 100, "accuracy" = c(rf_accuracy_4) * 100,"notes" = c("Trees = 500 and Only important variables"))
auc_df <- rbind(auc_df,auc_df_temp)

auc_df <- auc_df %>% 
  arrange(desc(auc))

auc_df
```

```{r target_company_row}
target_company_row_final
```

```{r run_against_target_company_rf}
class_probabilities_prediction <- predict(rf_model_4, newdata = target_company_row_final[, -1*c(target_var_index:target_var_index)], type = "prob")
class_probabilities_prediction
litigated_probablity <- class_probabilities_prediction$Yes
rf_litigated <- ifelse(litigated_probablity > 0.50, "Yes", "No")
rf_actual_litigated_num <- as.numeric(as.character(target_company_row_final$litigated))
rf_actual_litigated <- ifelse(rf_actual_litigated_num == 1, "Yes", "No")
rf_litigated
rf_actual_litigated
```

```{r run_against_target_company_gbm}
class_probabilities_prediction <- predict(gbm_model_1, newdata = target_company_row_final[, -1*c(target_var_index:target_var_index)], type = "prob")
litigated_probablity <- class_probabilities_prediction$Yes
gbm_litigated <- ifelse(litigated_probablity > 0.50, "Yes", "No")
gbm_actual_litigated_num <- as.numeric(as.character(target_company_row_final$litigated))
gbm_actual_litigated <- ifelse(rf_actual_litigated_num == 1, "Yes", "No")
gbm_litigated
gbm_actual_litigated
```

```{r run_against_target_company_glm}
class_probabilities_prediction <- predict(glm_model_3, newdata = target_company_row_final[, -1*c(target_var_index:target_var_index)], type = "prob")
litigated_probablity <- class_probabilities_prediction$Yes
glm_litigated <- ifelse(litigated_probablity > 0.50, "Yes", "No")
glm_actual_litigated_num <- as.numeric(as.character(target_company_row_final$litigated))
glm_actual_litigated <- ifelse(rf_actual_litigated_num == 1, "Yes", "No")
glm_litigated
glm_actual_litigated
```

```{r final_col_names_training}
nrow(training_subset_ds_final)
#summary(training_subset_ds_final$litigation_settlement)
#training_litigated <- training_subset_ds_final %>%
#  filter(!is.na(litigation_settlement))
#training_litigated  
```

```{r final_col_names_test}
nrow(test_subset_ds_final)
summary(test_subset_ds_final)
```



```{r final_col_names_test}
library(Rmisc)
settlement_training_ds <- training_subset_ds_final_settlement %>%
  filter(!is.na(litigation_settlement))

settlement_test_ds <- test_subset_ds_final_settlement %>%
  filter(!is.na(litigation_settlement))

settlement_ds <- rbind(settlement_training_ds,settlement_test_ds)
settlement_ds

settlement_ds_subset <- subset(settlement_ds, select = c(gvkey,tic, csho, prccd_m, prcod_m,litigated,litigation_settlement))
settlement_ds_subset$stk_price_m <- (settlement_ds_subset$prccd_m + settlement_ds_subset$prcod_m)/2
settlement_ds_subset$market_cap <- settlement_ds_subset$stk_price_m * settlement_ds_subset$csho * 1000000
settlement_ds_subset$settlement_ratio <- as.numeric(settlement_ds_subset$litigation_settlement/settlement_ds_subset$market_cap)
settlement_ds_subset

#CI(settlement_ds_subset$litigation_settlement, ci=0.90)
```

```{r}
settlement_gvkey <- settlement_ds_subset$gvkey
settlement_gvkey

fundamentals_ds_settlement <- fundamentals_ds_filter %>%
  filter(gvkey %in% settlement_gvkey)

fundamentals_ds_settlement_1 <- subset(fundamentals_ds_settlement, select = c(gvkey,tic, mkvalt))

fundamentals_ds_settlement_2 <- fundamentals_ds_settlement_1 %>%
  filter(!is.na(mkvalt)) 

fundamentals_ds_settlement_3 <- fundamentals_ds_settlement_2 %>%
   group_by(gvkey) %>%
   dplyr::summarize(
     market_value = mean(mkvalt)
   )
settlement_ds_subset$market_value <- 0
for (row in 1:nrow(settlement_ds_subset)){
  settlement_item_gvkey  <- as.integer(settlement_ds_subset[row, "gvkey"])
  fundamental_settlement_row <- fundamentals_ds_settlement_3 %>%
       filter(gvkey == settlement_item_gvkey)
  market_value  <- as.numeric(fundamental_settlement_row["market_value"])
  settlement_ds_subset$market_value[settlement_ds_subset$gvkey == settlement_item_gvkey] <- market_value * 1000000
}

settlement_ds_subset$market_value[settlement_ds_subset$tic == 'GRO'] <- as.numeric(72720449)

settlement_ds_subset$settlement_ratio_2 <- as.numeric(settlement_ds_subset$litigation_settlement/settlement_ds_subset$market_value)

settlement_ds_subset
#CI(settlement_ds_subset$litigation_settlement, ci=0.95)
settlement_ratio_mean <- mean(settlement_ds_subset$settlement_ratio_2)
settlement_ratio_medain <- median(settlement_ds_subset$settlement_ratio_2)
settlement_ratio_mean
settlement_ratio_medain

```



```{r}
target_company_ds_settlement <- fundamentals_ds_filter %>%
  filter(tic == 'SJM')
target_company_market_value <- mean(target_company_ds_settlement$mkvalt) * 1000000
target_company_settlement_amt <- target_company_market_value * settlement_ratio_medain
target_company_market_value
settlement_ratio_medain
target_company_settlement_amt
```

```{r}
settlement_1 <- settlement_ds_subset$litigation_settlement
#inference(settlement_1, est = "mean", type = "ci", method = "theoretical")
std_error <- function(x) sd(x)/sqrt(length(x))
std_error_settlement_amout <- std_error(settlement_1)
std_error_settlement_amout
settlement_amt_high <- target_company_settlement_amt + (1.96 * std_error_settlement_amout)
settlement_amt_low <- target_company_settlement_amt - (1.96 * std_error_settlement_amout) 
settlement_amt_high
settlement_amt_low
```

```{r}
settlement_1_sd = sd(settlement_1)
settlement_1_mean = mean(settlement_1)
settlement_1_sd
settlement_1_mean
z_value <- 1.96
n <- 11
#inference(settlement_1, est = "mean", type = "ci", method = "theoretical")
#inference(settlement_1, est = "median", type = "ci", method = "simulation")
```

