---
title: "Data-Due-Diligence"
author: "Amol Gote"
date: "6/6/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
library(tidyverse)
library(maps)
library(scales)
library(sf)
library(Hmisc)
library(dplyr)
library(ggplot2)
```
```{r loadData, message=FALSE}
customers <- read_csv("data/Customer_Dataset_File_Original.csv")
```
# Identify all missing values
```{r getAllMissingValueCount, message=FALSE}
sum(is.na(customers))
sapply(customers,function(x)sum(is.na(x)))
```

# Feature engineering steps
# Data Imputation for customers.
```{r imputeGender, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
print("Number of rows with Gender = NA before imputation")
sum(is.na(customers$Gender))
numberOfFemales <- customers %>% 
  filter(Gender=="Female") %>% 
  nrow()
print("Number of females before imputation")
numberOfFemales
numberOfMales <- customers %>% 
  filter(Gender=="Male") %>% 
  nrow()
count <- 0
print("Number of males before imputation")
numberOfMales
for(i in 1:nrow(customers)){
  if(is.na(customers$Gender[i])){
    #customers$Gender[i] <- "F"
    count <- count + 1
    if (count %% 2 == 0){
      customers$Gender[i] <- "Female"
    }
    else{
      customers$Gender[i] <- "Male"
    }
  }
}
print("Number of rows with Gender = NA after imputation")
sum(is.na(customers$Gender))
numberOfFemales <- customers %>% 
  filter(Gender=="Female") %>% 
  nrow()
print("Number of females after imputation")
numberOfFemales
numberOfMales <- customers %>% 
  filter(Gender=="Male") %>% 
  nrow()
print("Number of males after imputation")
numberOfMales
```

# Data Imputation for HouseholdSize.
```{r imputeHouseHold, message=FALSE}
print("Number of rows with HouseholdSize = NA before imputation")
sum(is.na(customers$HouseholdSize))
summary(customers$HouseholdSize)
customers$HouseholdSize[is.na(customers$HouseholdSize)] <- median(customers$HouseholdSize, na.rm = T)
print("Check for missing values for HouseholdSize")
sum(is.na(customers$HouseholdSize))
```

# Data Imputation for HomeOwner.
```{r imputeHoeOwnership, message=FALSE}
print("Number of rows with HomeOwner = NA before imputation")
sum(is.na(customers$HomeOwner))
summary(customers$HomeOwner)
get_mode <- function(x) {
  unique_x <- unique(x)
  mode <- unique_x[which.max(tabulate(match(x, unique_x)))]
  mode
}
mode_value = get_mode(customers$HomeOwner)
customers$HomeOwner[is.na(customers$HomeOwner)] <- mode_value
print("Check for missing values for HomeOwner")
sum(is.na(customers$HomeOwner))
ggplot(customers) + 
  geom_bar(aes(x = HomeOwner))
```

# Data Imputation for JobCategory
```{r imputeJobCategory, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
print("Number of rows with JobCategory = NA before imputation")
sum(is.na(customers$JobCategory))
summary(customers$JobCategory)
ggplot(customers) + 
  geom_bar(aes(x = JobCategory))
customers$JobCategory[is.na(customers$JobCategory)] <- "Misc"
print("Check for missing values for HomeOwner")
sum(is.na(customers$JobCategory))
ggplot(customers) + 
  geom_bar(aes(x = JobCategory))
```

# Remove the dollar sign from household income
```{r cleanAmountVariables, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
customers <- customers %>%
  replace(.=="#NULL!", NA) # replace with NA
customers$HHIncome = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$HHIncome)))) 
customers$VoiceLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$VoiceLastMonth)))) 
customers$VoiceOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$VoiceOverTenure)))) 
customers$CardSpendMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$CardSpendMonth)))) 
customers$EquipmentLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$EquipmentLastMonth)))) 
customers$EquipmentOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$EquipmentOverTenure)))) 
customers$DataLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$DataLastMonth)))) 
customers$DataOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$DataOverTenure)))) 
sapply(customers,function(x)sum(is.na(x)))
ncol(customers)
```


# Remove the dollar sign from CarValue
```{r imputeCarValue, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
customers$CarValue = gsub("\\ ", "", customers$CarValue) 
summary(customers$CarValue)
carValueLessThanZero <- customers %>% 
  filter(CarValue =="$(1,000.00)") %>% 
  nrow()
customers$CarValue[customers$CarValue =="$(1,000.00)"] <- 0
customers$CarValue = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$CarValue)))) 
customers$CarValue = as.numeric(customers$CarValue)
carValueLessThanZero <- customers %>% 
  filter(is.na(customers$CarValue)) %>%
  nrow()
  
print("Number of records with CarValue = NA")
carValueLessThanZero
carOwnership <- customers %>% 
  filter(customers$CarOwnership == "-1") %>%
  nrow()
print("Number of records with CarOwnership == -1")
carOwnership
carBrand <- customers %>% 
  filter(customers$CarBrand == "-1") %>%
  nrow()
print("Number of records with CarBrand == -1")
carBrand
customers$CarOwnership[customers$CarOwnership =="-1"] <- NA
customers$CarBrand[customers$CarBrand =="-1"] <- NA
```

# Impute Commute Time, there are 2 missing values
```{r imputeCommuteTime, message=FALSE}
customers$CommuteTime = as.numeric(customers$CommuteTime)
customers$CommuteTime[is.na(customers$CommuteTime)] <- mean(customers$CommuteTime, na.rm = T)
ggplot(customers) + 
  geom_bar(aes(x = CommuteTime))
```
#Address Missing values for EquipmentLastMonth, DataLastMonth, EquipmentOverTenure,DataOverTenure
```{r addressMissingValies, message=FALSE}
customers$EquipmentLastMonth[is.na(customers$EquipmentLastMonth)] <- 0
customers$DataLastMonth[is.na(customers$DataLastMonth)] <- 0
customers$EquipmentOverTenure[is.na(customers$EquipmentOverTenure)] <- 0
customers$DataOverTenure[is.na(customers$DataOverTenure)] <- 0
customers$PhoneCoTenure <- ifelse(customers$PhoneCoTenure == 0, 1, customers$PhoneCoTenure)
customers$VoiceOverTenure <- ifelse(is.na(customers$VoiceOverTenure), customers$VoiceLastMonth, customers$VoiceOverTenure)
```

# Final Summary of customer dataset
```{r finalSummary, message=FALSE}
summary(customers)
```


# Adding of additional variables 
```{r additionalVaribales, message=FALSE}
customers$LastMonthTotalValue <- customers$VoiceLastMonth + customers$EquipmentLastMonth + customers$DataLastMonth
customers$OverTenureTotalValue <- customers$VoiceOverTenure + customers$EquipmentOverTenure + customers$DataOverTenure
customers$TotalDebt <- customers$CreditDebt + customers$OtherDebt
customers$AverageMonthlyRevenue <- customers$OverTenureTotalValue/customers$PhoneCoTenure
```

# Create a dataset for analysis with required variables only
```{r filteredColumnData, message=FALSE}
customersForAnalysis <- customers[,c("CustomerID","Gender","Age","EducationYears","JobCategory","EmploymentLength","HHIncome","DebtToIncomeRatio","CreditDebt","OtherDebt","MaritalStatus","HouseholdSize","HomeOwner","CommuteTime","PhoneCoTenure","VoiceLastMonth","VoiceOverTenure","EquipmentRental","EquipmentLastMonth","EquipmentOverTenure","WirelessData","DataLastMonth","DataOverTenure","Multiline","EBilling", "LastMonthTotalValue", "OverTenureTotalValue", "TotalDebt", "AverageMonthlyRevenue")]
summary(customersForAnalysis)
```
#Statistical analysis for Age variable
```{r agePlots, message=FALSE}
ggplot(data = customersForAnalysis,
       aes(x = Gender, y = AverageMonthlyRevenue)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom="point",size=3) +
  stat_summary(fun.y = median, geom="point",size=3) 

customersForAnalysis%>%
group_by(Gender)%>% 
summarise(Mean=mean(AverageMonthlyRevenue), Median=median(AverageMonthlyRevenue))
```
# Age distribution and monthly revenue genegrate by age groups
```{r ageDistribution, message=FALSE}
ggplot(customersForAnalysis, aes(x=Age)) +
    geom_histogram(binwidth=.5, colour="black", fill="white") + 
    labs( title= "Age frequency distribution", x="Age in years", y = "Number of customers")

lbls <- c('0-10','10-20','20-30', '30-40','40-50','50-60','60-70', '70-80', '80-90', '90-100')
groupedData <- customersForAnalysis %>% 
     group_by(ageGroup = cut(Age, breaks= seq(0, 100, by = 10), 
                                right = TRUE, include.lowest = TRUE, labels = lbls) ) %>% 
     summarise(totalMonthlyValue= sum(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = ageGroup, y = totalMonthlyValue)) + 
  geom_bar(stat="identity", width=0.5, fill = "steelblue") +
  labs( title= "Monthly revenue generated by age groups", x="Age groups", y = "Monthly revenue ($)")
```
#Statistical analysis for total debt
```{r totalDebt, message=FALSE}
lbls <- c('0-5','5-10','10-15', '15-20','20-25','25-30','30-35', '35-40', '40-45', '45-50')
groupedData <- customersForAnalysis %>% 
     group_by(debtGroup = cut(TotalDebt, breaks= seq(0, 50, by = 5), 
                                right = TRUE, include.lowest = TRUE, labels = lbls) ) %>% 
     summarise(totalMonthlyValue= sum(AverageMonthlyRevenue)) 


ggplot(groupedData, aes(x = debtGroup, y = totalMonthlyValue)) + 
  geom_bar(stat="identity", width=0.5, fill = "steelblue") +
  labs( title= "Monthly revenue generated by total debt(Buckets of 5)", x="Total debt range", y = "Monthly revenue ($)")
```
#Statistical analysis for Household size
```{r hoseHoldSize, message=FALSE}
groupedData <- customersForAnalysis%>%
  group_by(HouseholdSize)%>% 
  summarise(Mean=mean(AverageMonthlyRevenue), Median=median(AverageMonthlyRevenue), totalMonthlyRevenue= sum(AverageMonthlyRevenue))
groupedData
```

##Household size > 1
```{r householdSizeGreaterThan1, message=FALSE }
groupedData <- customersForAnalysis%>%
  filter(HouseholdSize > 1) %>%
  filter(Multiline == "No") %>%
  summarise(totalMonthlyRevenue= sum(AverageMonthlyRevenue))
cat(sprintf("Total monthly revenue for household size > 1 and with no multi-line:: %s", groupedData$totalMonthlyRevenue))
groupedData <- customersForAnalysis%>%
  filter(HouseholdSize > 1) %>%
  filter(Multiline == "No")
numberOfCustomers <- count(groupedData)
cat(sprintf("\nTotal monthly revenue for household size > 1 and with no multi-line: %s", numberOfCustomers))
```
#Statistical analysis for Education
```{r educationInYears}
groupedData <- customersForAnalysis %>% 
     group_by(EducationYears) %>% 
     summarise(totalMonthlyValue= sum(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = EducationYears, y = totalMonthlyValue)) + 
  geom_bar(stat="identity", width=0.5, fill = "steelblue") +
  labs( title= "Monthly revenue generated by Education in years(Buckets of 5)", x="Education (in years)", y = "Monthly revenue ($)")
```
#Statistical analysis for household income
```{r houseHoldIncome, message=FALSE}
ggplot(customersForAnalysis, aes(x = HHIncome, y = AverageMonthlyRevenue)) + 
  geom_point() +
  labs( title= "Distribution of household income and average monthly revenue", x="House hold income", y = "Monthly revenue ($)")

lbls <- c('0-10K','10-20K','20-30K', '30-40K','40-50K','50-60K','60-70K', '70-80K', '80-90K', '90-100K', '100-110K', '110-120K', '120-130K', '130-140K', '140-150K')

groupedData <- customersForAnalysis %>% 
     group_by(hhIncomeGroup = cut(HHIncome, breaks= seq(0, 150000, by = 10000), 
                                right = TRUE, include.lowest = TRUE, labels = lbls) ) %>% 
     summarise(totalMonthlyValue= sum(AverageMonthlyRevenue))
groupedData

ggplot(groupedData, aes(x = hhIncomeGroup, y = totalMonthlyValue)) + 
  geom_bar(stat="identity", width=0.5, fill = "steelblue") +
  labs( title= "Monthly revenue generated by household income(Buckets of $10K)", x="Household income", y = "Monthly revenue ($)") +
  theme(axis.text.x = element_text(angle = 45))
```

#Statistical analysis for Job category
```{r jobCategory, message=FALSE}
groupedData <- customersForAnalysis%>%
group_by(JobCategory)%>% 
summarise(totalMonthlyValue = sum(AverageMonthlyRevenue))

groupedData

ggplot(groupedData, aes(x = JobCategory, y = totalMonthlyValue)) + 
  geom_bar(stat="identity", width=0.5, fill = "steelblue") +
  labs( title= "Monthly revenue generated by job categories", x="Job category", y = "Monthly revenue ($)") 
```

#Statistical analysis for home ownership
```{r homeOwnership, message=FALSE}
ggplot(data = customersForAnalysis, aes(x = HomeOwner, y = AverageMonthlyRevenue, group=HomeOwner)) +
  geom_boxplot() +
  labs( title= "Homeownership and monthly revenue generated", x="Home ownership", y = "Monthly revenue ($)") 

customersForAnalysis%>%
group_by(HomeOwner)%>% 
summarise(SUm = sum(AverageMonthlyRevenue), Mean=mean(AverageMonthlyRevenue), Median=median(AverageMonthlyRevenue))
```

#Statistical analysis for Debt to income ratio
```{r debtToIncomeRatio, message=FALSE}
lbls <- c('0-10','10-20','20-30', '30-40','40-50','50-60','60-70', '70-80', '80-90', '90-100')
groupedData <- customersForAnalysis %>% 
     group_by(debtToIncomeRatioGroup = cut(DebtToIncomeRatio, breaks= seq(0, 100, by = 10), 
                                right = TRUE, include.lowest = TRUE, labels = lbls) ) %>% 
     summarise(totalMonthlyValue= sum(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = debtToIncomeRatioGroup, y = totalMonthlyValue)) + 
  geom_bar(stat="identity", width=0.5, fill = "steelblue") +
  labs( title= "Monthly revenue generated Debt to income ration (Buckets of 10)", x="Debt to income ratio (DTI)", y = "Monthly revenue ($)") 
```

#Export the dataset used for analysis
```{r exportDateset, message=FALSE}
write_csv(customersForAnalysis,"data/Customer_Dataset_UPDATED.csv")
```
