---
title: "Predictive_Modeling_Week_1_Excercise_1"
author: "Amol Gote"
date: "7/11/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
library(tidyverse)
library(plotly)
library(gpairs) 
library (MASS)
library(ggplot2)
```
# Read File
```{r problem_8a, message=FALSE}
college <- read.csv("data/College.csv")
#college

```
# Read Fix  College name column 
```{r problem_8b, message=FALSE}
#rownames(college) = college[,1]
college_orig <- college
college =college [,-1]
#fix(college)
#college
```
# Summary
```{r problem_8c_i, message=FALSE}
summary(college)
```

# pairs() function
```{r problem_8c_ii, message=FALSE}
pairs(college[,1:10])
pairs(∼ Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + 
         P.Undergrad + Outstate + Room.Board + Books, college)
```

```{r problem_8c_ii_explicit_variables, message=FALSE}
pairs(college[,1:10])
pairs(∼ Apps + Accept + Enroll + Top10perc + Top25perc + F.Undergrad + 
         P.Undergrad + Outstate + Room.Board + Books, college)
```

# plot function
```{r problem_8c_iii, message=FALSE}
plot(college$Private, college$Outstate, 
     xlab = "Private University", ylab ="Out of State", main = "Outstate versus Private")
```

# Elite
```{r problem_8c_iv, message=FALSE}
Elite = rep("No",nrow(college))
Elite[college$Top10perc > 50]= "Yes"
Elite = as.factor (Elite)
college = data.frame(college, Elite)
summary(college)
summary(college$Elite)
plot(college$Elite, college$Outstate, 
     xlab = "Elite University", ylab ="Out of State", main = "Outstate versus Private")
```

# Histograms
```{r problem_8c_v, message=FALSE}
par(mfrow=c(3,2))
hist(college$Apps, col = 2, xlab = "Number of applications", ylab = "Count", main = "Distribution of applications")
hist(college$Accept, col = 4, xlab = "Number of Accepted applicatons", ylab = "Count", main = "Distribution of college acceptance")
hist(college$Enroll, col = 4, xlab = "Number of Enrollments", ylab = "Count", main = "Distribution of college enrollment")
hist(college$Grad.Rate, col = 5, xlab = "Graduation Rate", ylab = "Count", main = "Distribution of graduation rates")
hist(college$PhD, col = 6, xlab = "PhD", ylab = "Count", main = "Distribution of Phd's")
hist(college$Books, col = 7, xlab = "Books", ylab = "Count", main = "Distribution of Books cost")
```

# Continue exploring the data, and provide a brief summary of what you discover.
```{r problem_8c_vi, message=FALSE}
college$rejected <- college$Apps - college$Accept
hist(college$rejected, col = 2, xlab = "Number of rejected", ylab = "Count", main = "Distribution of applications")
summary(college$rejected)

colleges_rejected_5000 <- college %>%
  filter(college$rejected > 5000)
nrow(colleges_rejected_5000)

colleges_rejected_20000 <- college %>%
  filter(college$rejected > 20000)
colleges_rejected_20000

colleges_with_high_reject <- college_orig %>%
  filter(Apps == 48094)
colleges_with_high_reject

summary(college$Outstate)

hist(college$Outstate, col = 3, xlab = "Out-of-state tuition", ylab = "Count", main = "Distribution of Out-of-state tuition cost")

```

1. Difference between number of applications and accepted indicates how many applications have been rejected. 
2. Average rejected applications is 928 and median is 326. 
3. There are 24 colleges with number of applications rejected > 5000. 
5. "Rutgers at New Brunswick" college has the highest rejected applications as well as highest number of applications which is 48090. 
6. College out state tuition is evenly distributed with mean and median around 10000 

# Problem 10
# Boston dataset info
```{r problem_10a, message=FALSE}
#?Boston
#Boston
nrow(Boston)
ncol(Boston)
```

1. Dataset has 506 rows 
2. Dataset has 14 columns.
3. Each column represents variable for which an observation was made in 506 neighborhoods of Boston.These variables include 
    a. crim - per capita crime rate by town. 
    b. zn - proportion of residential land zoned for lots over 25,000 sq.ft. 
    c. indus - proportion of non-retail business acres per town. 
    d. chas - Charles River dummy variable (= 1 if tract bounds river; 0 otherwise). 
    e. nox - nitrogen oxides concentration (parts per 10 million). 
    f. rm - average number of rooms per dwelling. 
    g. age - proportion of owner-occupied units built prior to 1940. 
    h. dis - weighted mean of distances to five Boston employment centres. 
    i. rad - index of accessibility to radial highways. 
    j. tax - full-value property-tax rate per \$10,000. 
    k. ptratio - pupil-teacher ratio by town. 
    l. black - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town. 
    n. lstat - lower status of the population (percent). 
    o. medv - median value of owner-occupied homes in \$1000s. 
  
  
# Pairwise scatterplots of the predictors
```{r problem_10b, message=FALSE}
pairs(∼ crim + zn + 
       + rm
       + age 
       + dis 
       + tax 
       + ptratio
       + black
       + lstat
       + medv, Boston
       )
pairs(crim ~ age + tax + black + medv + rad, Boston)
pairs(nox ~ zn, Boston)
pairs(medv ~ tax + lstat  + nox + rm, Boston)
```

1. Where Proportion of residential land zoned is less the nox levels have increased. 
2. As lsat (lower status of the population) increases then medv(median value of owner-occupied homes) decreases.  
3. As nox (lower status of the population) increases then medv(median value of owner-occupied homes) decreases. 
4. As rm (average number of rooms per dwelling) increases then medv(median value of owner-occupied homes) increases. 

# predictors associated with per capita crime rate?
```{r problem_10c, message=FALSE}
pairs(crim ~ rad, Boston)
pairs(crim ~ age, Boston)
pairs(crim ~ medv, Boston)
pairs(crim ~ tax,  Boston)
```

1. A higher values of rad (index of accessibility to radial highways) contain the highest level of crime rate. 
2. As the age (proportion of owner-occupied units built prior to 1940) has increased the crime rate has increased. 
3. crime rate is higher in the lower median value of owner-occupied homes. 
4. As tax rate increase, crime rate increases. 

# suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios
```{r problem_10d_1, message=FALSE}
summary(Boston$crim)
crimeGreateThan20 <- Boston$crim[Boston$crim>20]
crimeGreateThan30 <- Boston$crim[Boston$crim>30]
length(crimeGreateThan20)
length(crimeGreateThan30)
par(mfrow=c(3,2))
hist(Boston$crim, col = 3, xlab = "Crime rate", ylab = "Suburbs", main = "Distribution of Crime rate")
highCrime <- Boston$crim[Boston$crim>20]
hist(highCrime, col = 3, xlab = "Crime rate", ylab = "Suburbs", main = "Distribution of Crime rate")
highCrime <- Boston$crim[Boston$crim>30]
hist(highCrime, col = 3, xlab = "Crime rate", ylab = "Suburbs", main = "Distribution of Crime rate")
```
```{r problem_10d_2, message=FALSE}
par(mfrow=c(1,2))
hist(Boston$tax, col = 3, xlab = "Tax", ylab = "# Suburbs", main = "Distribution of tax")
hist(Boston$ptratio, col = 8, xlab = "Pupil teacher ratios", ylab = "# Suburbs", main = "Distribution of Pupil-teacher ratios")
summary(Boston$tax)
summary(Boston$ptratio)
length(Boston$crim[Boston$tax>600])
length(Boston$crim[Boston$ptratio>20])
```

1. Most cities have less crime rate with average crime rate = 3.6. 
2. There are 18 suburbs with crime rate greater than 20 and 8 suburbs with crime rate greater than 30. 
3. Some suburbs have the highest tax with value greater than 600, there are 137 suburbs like these. These suburbs are beyond the 3rd quartile values. 
4. Some suburbs have the highest Pupil-teacher ratios with value greater than 20, there are 201 suburbs like these. These suburbs are beyond the 3rd quartile values. 

# Suburbs in this data set bound the Charles river
```{r problem_10e, message=FALSE}
suburbs_bound_to_river <- Boston %>%
  filter(chas == 1)
nrow(suburbs_bound_to_river)
```

1. Number of Suburbs bound to the Charles river is 35. 

# What is the median pupil-teacher ratio among the towns in this data set?
```{r problem_10f, message=FALSE}
median(Boston$ptratio)
```

1. Median pupil-teacher ratio is 19.05. 


# Which suburb of Boston has lowest median value of owneroccupied homes?
```{r problem_10g_1, message=FALSE}
lowestMedianValue <- Boston %>%
  filter(medv == min(Boston$medv))
lowestMedianValue
```

1. There are 2 suburbs with with lowest median value of 5.

```{r compare_summary_function, message=FALSE}
compare_summary_dataset <- function(dataset_orig,dataset_target) {
  sum1 <- apply(dataset_orig,2,summary) %>% data.frame() 
  sum2 <- apply(dataset_target,2,summary) %>% data.frame() 

  names(sum1) <- paste0(names(sum1),"1")
  names(sum2) <- paste0(names(sum2),"2")

  final <- cbind(sum1,sum2)

  final1 <- t(final) 

  final2 <- final1[order(row.names(final1)), ]

  final_1 <- t(final2) %>% data.frame()
  final_1
}
```

# What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors? Comment on your findings.
```{r problem_10g_2, message=FALSE}
t(compare_summary_dataset(Boston,lowestMedianValue))
```

1. Age of lowest median value of owner-occupied homes is higher compared to overall dataset. 
2. Proportion of blacks is same as that of median of overall dataset, mean of these 2 suburbs is on higher side. 
3. Crime rate is on the higher side; 2 suburbs have crime rate of 38 and 67 with average of 53 which is considerably higher than the overall dataset. 
4. Lower status of the population (percent) is greater than mean and median of overall dataset. 
5. nox (nitrogen oxides concentration (parts per 10 million)) is on higher side. 
6. ptratio (pupil-teacher ratio) falls in that tower of > 20 in the above histogram.  
7. Tax is on the higher side compared to overall dataset and it falls in the tower > 600 histogram. 


# Suburbs average more than seven rooms per dwelling
```{r problem_10h_1, message=FALSE}
dwelling_7 <- Boston %>%
  filter(rm > 7)
nrow(dwelling_7)
```

# Suburbs average more than eight rooms per dwelling
```{r problem_10h_2, message=FALSE}
dwelling_8 <- Boston %>%
  filter(rm > 8)
nrow(dwelling_8)
summary(dwelling_8)
```

1. Number of suburbs with more than seven rooms per dwelling is 64. 
2. Number of suburbs with more than eight rooms per dwelling is 13. 

# Suburbs that average more than eight rooms per dwelling compared to overall dataset.
```{r problem_10h_3, message=FALSE}
t(compare_summary_dataset(Boston,dwelling_8))
```

1. Crime rate for suburbs with more than eight rooms per dwelling is on lower side compared to overall dataset. 
2. lstat (lower status of the population (percent)) is lower than the overall population. 
3. medv (median value of owner-occupied homes) is on higher side than overall dataset. 
4. tax (full-value property-tax rate per $10,000) mean is lower than that of the overall dataset. 



