---
title: "Week8-Lab VII"
author: "Aamol Gote"
date: "10/18/2019"
output:
  html_document: default
  word_document: default
---

```{r}
download.file("http://www.openintro.org/stat/data/mlb11.RData", destfile = "mlb11.RData")
load("mlb11.RData")
```

```{r}
nrow(mlb11)
```

**Excercise 1**

What type of plot would you use to display the relationship between runs and one of the other numerical variables? Plot this relationship using the variable at_bats as the predictor. Does the relationship look linear? If you knew a team’s at_bats, would you be comfortable using a linear model to predict the number of runs?

Scatterplot would be used to display the relationship between runs and one of the other Numerical variables

```{r}
plot(mlb11$at_bats, mlb11$runs, main = "Relationship between runs and at_bats", xlab = "at_bats", ylab = "Runs")
```

Relationship between runs and at_bats is linear, but relationship is not strong relationship and it seems to be moderate one.

We can quantify the relationship between runs and at_bats using correlation coefficient

```{r}
cor(mlb11$at_bats, mlb11$runs)
```


Team’s at_bats can be used in a linear model to predict the number of runs given that we could see some linear trend in relationship. But, the relationship is not strong enough, so the prediction may not be of desired accuracy


**Excercise 2**

Looking at your plot from the previous exercise, describe the relationship between these two variables. Make sure to discuss the form, direction, and strength of the relationship as well as any unusual observations.

1. Relationship between runs and at_bats is positively correlated, as at_bats increases runs also increases.
2. Relationship is not strong but moderatly stronger.
3. There are few outliars like at_bats > 5650 as well runs near 850
4. There’s no curvature to the plot, so this is a linear relationship.


**Excercise 3**

Using plot_ss, choose a line that does a good job of minimizing the sum of squares. Run the function several times. What was the smallest sum of squares that you got? How does it compare to your neighbors?


```{r}
plot_ss(x = mlb11$at_bats, y = mlb11$runs)
```

Smallest Sum of squares: 123721.9

The neighboring value deviate from the smallest value by around 4000 - 5000.

==============================================

```{r}
m1 <- lm(runs ~ at_bats, data = mlb11)
summary(m1)
```


**Excercise 4**

Fit a new model that uses homeruns to predict runs. Using the estimates from the R output, write the equation of the regression line. What does the slope tell us in the context of the relationship between success of a team and its home runs?

```{r}
m1 <- lm(runs ~ homeruns, data = mlb11)
summary(m1)
```

Equation of regression line

y = b0 + b1x

runs = 415.2389 + 1.8345 * homeruns

In the context of the relationship between success of a team and its home runs slope tells that for every home run, number of total runs will also increase 1.8345.

Slope = 1.8345

```{r}
cor(mlb11$runs, mlb11$homeruns)
```

Positive relationship with a correlation coefficient of 0.7915577, which is on stronger side (relatively stronger).



```{r}
plot(mlb11$runs ~ mlb11$at_bats)
abline(m1)
```

**Excercise 5**

If a team manager saw the least squares regression line and not the actual data, how many runs would he or she predict for a team with 5,578 at-bats? Is this an overestimate or an underestimate, and by how much? In other words, what is the residual for this prediction?

```{r}
m1 <- lm(runs ~ at_bats, data = mlb11)
summary(m1)
```

Least Square Regression line for runs vs at_bats

y = b0 + b1x

b0 = -2789.2429
b1 = 0.6305;

x = 5578 (at_bats)

y = -2789.2429 + 0.6305 * (5578)

y =727.6861

Predicted runs (y) = 728

Estimated number of runs for 5578 at bats based on the linear regression formula above is 728.

Based on below data there are no teams with at_bats = 5578, nearest match is at_bats = 5579 for **Philadelphia Phillies** with runs **713**. So going by this there is an overestimate of 728 - 713 = 15 runs.

Residual of the prediction is 15


```{r}
mlb11[order(mlb11$runs,mlb11$at_bats),]

```

=================

```{r}
plot(m1$residuals ~ mlb11$at_bats)
abline(h = 0, lty = 3)
```

**Excercise 6**

Is there any apparent pattern in the residuals plot? What does this indicate about the linearity of the relationship between runs and at-bats?

There is no apparent pattern in the residual plot. 

Distribution is scattered all over. It does indicate linear relationship between runs and at_bats.


=============================================================

```{r}
hist(m1$residuals)
```

```{r}
qqnorm(m1$residuals)
qqline(m1$residuals)
```

**Excercise 7**

Based on the histogram and the normal probability plot, does the nearly normal residuals condition appear to be met?

Histogram is right skewed.

Nearly normal residual condition appears to be satisfied.


**Excercise 8**

Based on the plot in (1), does the constant variability condition appear to be met?

Constant variability condition appears to be satisfied barring some out-liars.


=============================================================


**On Your Own**


1. Choose another traditional variable from mlb11 that you think might be a good predictor of runs. Produce a scatterplot of the two variables and fit a linear model. At a glance, does there seem to be a linear relationship?

Another variable considered bat_avg (Batting average), reason for choosing the same is higher the batting average higher number of runs.

```{r}
plot(mlb11$bat_avg, mlb11$runs, main = "Relationship between runs and Batting Average", xlab = "at_bats", ylab = "Runs")
m3 <- lm(runs ~ bat_avg, data = mlb11)
abline(m3)
```

```{r}
cor(mlb11$runs, mlb11$bat_avg)
```
```{r}
summary(m3)
```

```{r}
summary(m1)
```

Least Square Regression line for runs vs bat_avg

y = b0 + b1x

b0 = -642.8
b1 = 5242.2

y = -642.8 + (5242.2)*x

Relationship between runs vs bat_avg is positive, linear and relatively strong.



2. How does this relationship compare to the relationship between runs and at_bats? Use the R2 values from the two model summaries to compare. Does your variable seem to predict runs better than at_bats? How can you tell?

Relationship between runs and bat_avg seems to be more stronger than runs and at_bats. 

R2 value for runs and bat_avg = 0.6561

R2 value for runs and at_bats = 0.3729

The variable “bat_avg” is a much better predictor of “runs” than the variable “at_bats” as is indcated by the R2 value.



3. Now that you can summarize the linear relationship between two variables, investigate the relationships between runs and each of the other five traditional variables. Which variable best predicts runs? Support your conclusion using the graphical and numerical methods we’ve discussed (for the sake of conciseness, only include output for the best variable, not all five). 


```{r}
hits <- lm(runs ~ hits, data = mlb11)
summary(hits)
```

```{r}
homeruns <- lm(runs ~ homeruns, data = mlb11)
summary(homeruns)
```

```{r}
bat_avg <- lm(runs ~ bat_avg, data = mlb11)
summary(bat_avg)
```

```{r}
strikeouts <- lm(runs ~ strikeouts, data = mlb11)
summary(strikeouts)
```

```{r}
stolen_bases <- lm(runs ~ stolen_bases, data = mlb11)
summary(stolen_bases)
```

```{r}
wins <- lm(runs ~ wins, data = mlb11)
summary(wins)
```

```{r}
cor(mlb11$runs, mlb11$hits)
cor(mlb11$runs, mlb11$homeruns)
cor(mlb11$runs, mlb11$bat_avg)
cor(mlb11$runs, mlb11$strikeouts)
cor(mlb11$runs, mlb11$stolen_bases)
cor(mlb11$runs, mlb11$wins)
```

Comparing R2 values and correlation coefficient for remaining variables

hits             => Multiple R-squared:  0.6419      

homeruns         => Multiple R-squared:  0.6266

bat_avg          => Multiple R-squared:  0.6561

strikeouts       => Multiple R-squared:  0.1694

stolen_bases     => Multiple R-squared:  0.002914

wins             => Multiple R-squared:  0.361

hits             => Correlation Coefficent:  0.8012108      

homeruns         => Correlation Coefficent:  0.7915577

bat_avg          => Correlation Coefficent:  0.8099859

strikeouts       => Correlation Coefficent:  -0.4115312

stolen_bases     => Correlation Coefficent:  0.05398141

wins             => Correlation Coefficent:  0.6008088


After running the summary statistics for all the variables, the variable which best predicts the runs based on R2 happened to be bat_avg. 

Correlation coeficient of bat_avg seems to be much better than other variables.

```{r}
plot(mlb11$runs ~ mlb11$bat_avg, main = "Relationship between Runs and Batting Avg", xlab = "Batting Avg", ylab = "Runs")
bat_avg <- lm(runs ~ bat_avg, data = mlb11)
abline(bat_avg)
```

```{r}
hist(bat_avg$residuals)
```


```{r}
qqnorm(bat_avg$residuals)
qqline(bat_avg$residuals)
```



4. Now examine the three newer variables. These are the statistics used by the author of Moneyball to predict a teams success. In general, are they more or less effective at predicting runs that the old variables? Explain using appropriate graphical and numerical evidence. Of all ten variables we’ve analyzed, which seems to be the best predictor of runs? Using the limited (or not so limited) information you know about these baseball statistics, does your result make sense?

Three new variables
  new_onbase
  new_slug
  new_obs

```{r}
new_onbase <- lm(runs ~ new_onbase, data = mlb11)
summary(new_onbase)
```

```{r}
new_slug <- lm(runs ~ new_slug, data = mlb11)
summary(new_slug)
```

```{r}
new_obs <- lm(runs ~ new_obs, data = mlb11)
summary(new_obs)
```
```{r}
cor(mlb11$runs, mlb11$new_onbase)
cor(mlb11$runs, mlb11$new_slug)
cor(mlb11$runs, mlb11$new_obs)
```

new_onbase       => Multiple R-squared:  0.8491 

new_slug         => Multiple R-squared:  0.8969

new_obs          => Multiple R-squared:  0.6561

new_onbase       => Correlation Coefficent:  0.9214691    

new_slug         => Correlation Coefficent:  0.9470324

new_obs          => Correlation Coefficent:  0.9669163


Earlier Best prediction model was based on bats_avg based on tradition variables

bat_avg          => Multiple R-squared:  0.6561

bat_avg          => Correlation Coefficent:  0.8099859


The newer variable seems to predict runs better than old variables. The R2 value for newer variables are higher than that of old variable. Given that the newer variables represent more advanced statistics of baseball it does make sense that they are better predictor of runs. So in summary result does make sense.

Of the new three variables the relationship between runs and new_obs variable has the highest R2. It appears to be best predictor of runs.

Below are the plots for all three new varaibles

```{r}
plot(mlb11$runs ~ mlb11$new_onbase)
abline(new_onbase)
```
```{r}
plot(mlb11$runs ~ mlb11$new_slug)
abline(new_slug)
```
```{r}
plot(mlb11$runs ~ mlb11$new_obs)
abline(new_obs)
```

5. Check the model diagnostics for the regression model with the variable you decided was the best predictor for runs.

new_obs is the best predictor.

Linearity

```{r}
plot(new_obs$residuals ~ mlb11$new_obs)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

The relationship looks linear based on a residual plot as the variability of residuals is approximately constant across the distribution but does not indicate any curvatures or any indication of non-normality.



Normal residuals:

```{r}
hist(new_obs$residuals)
```

The residuals are approximately normaly distributed.


Constant variability

```{r}
qqnorm(new_obs$residuals)
qqline(new_obs$residuals)
```

The least squares line remains roughly constant so the condition constant variability has been met
