---
title: "Data-Due-Diligence"
author: "Amol Gote"
date: "6/6/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(maps)
library(scales)
library(sf)
library(Hmisc)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

```{r include = FALSE}
#knitr::opts_chunk$set(echo=FALSE)
```

```{r loadData, message=FALSE}
customers <- read_csv("data/Customer_Dataset_File_Original.csv")
```
# Feature engineering steps
# Data Imputation for customers.
```{r imputeGender, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
for(i in 1:nrow(customers)){
  if(is.na(customers$Gender[i])){
    #customers$Gender[i] <- "F"
    count <- count + 1
    if (count %% 2 == 0){
      customers$Gender[i] <- "Female"
    }
    else{
      customers$Gender[i] <- "Male"
    }
  }
}
```

# Data Imputation for HouseholdSize.
```{r imputeHouseHold, message=FALSE}
customers$HouseholdSize[is.na(customers$HouseholdSize)] <- median(customers$HouseholdSize, na.rm = T)
```

# Data Imputation for HomeOwner.
```{r imputeHoeOwnership, message=FALSE}
get_mode <- function(x) {
  unique_x <- unique(x)
  mode <- unique_x[which.max(tabulate(match(x, unique_x)))]
  mode
}
mode_value = get_mode(customers$HomeOwner)
customers$HomeOwner[is.na(customers$HomeOwner)] <- mode_value
```

# Data Imputation for JobCategory
```{r imputeJobCategory, message=FALSE}
customers$JobCategory[is.na(customers$JobCategory)] <- "Misc"
```

# Remove the dollar sign from household income
```{r cleanAmountVariables, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
customers <- customers %>%
  replace(.=="#NULL!", NA) # replace with NA
customers$HHIncome = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$HHIncome)))) 
customers$VoiceLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$VoiceLastMonth)))) 
customers$VoiceOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$VoiceOverTenure)))) 
customers$CardSpendMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$CardSpendMonth)))) 
customers$EquipmentLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$EquipmentLastMonth)))) 
customers$EquipmentOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$EquipmentOverTenure)))) 
customers$DataLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$DataLastMonth)))) 
customers$DataOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$DataOverTenure)))) 
sapply(customers,function(x)sum(is.na(x)))
```


# Remove the dollar sign from CarValue
```{r imputeCarValue, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
customers$CarValue = gsub("\\ ", "", customers$CarValue) 
summary(customers$CarValue)
carValueLessThanZero <- customers %>% 
  filter(CarValue =="$(1,000.00)") %>% 
  nrow()
customers$CarValue[customers$CarValue =="$(1,000.00)"] <- 0
customers$CarValue = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$CarValue)))) 
customers$CarValue = as.numeric(customers$CarValue)
carValueLessThanZero <- customers %>% 
  filter(is.na(customers$CarValue)) %>%
  nrow()
  
carOwnership <- customers %>% 
  filter(customers$CarOwnership == "-1") %>%
  nrow()
carBrand <- customers %>% 
  filter(customers$CarBrand == "-1") %>%
  nrow()
customers$CarOwnership[customers$CarOwnership =="-1"] <- NA
customers$CarBrand[customers$CarBrand =="-1"] <- NA
```

# Impute Commute Time, there are 2 missing values
```{r imputeCommuteTime, message=FALSE}
customers$CommuteTime = as.numeric(customers$CommuteTime)
customers$CommuteTime[is.na(customers$CommuteTime)] <- mean(customers$CommuteTime, na.rm = T)
```
#Address Missing values for EquipmentLastMonth, DataLastMonth, EquipmentOverTenure,DataOverTenure
```{r addressMissingValies, message=FALSE}
customers$EquipmentLastMonth[is.na(customers$EquipmentLastMonth)] <- 0
customers$DataLastMonth[is.na(customers$DataLastMonth)] <- 0
customers$EquipmentOverTenure[is.na(customers$EquipmentOverTenure)] <- 0
customers$DataOverTenure[is.na(customers$DataOverTenure)] <- 0
customers$PhoneCoTenure <- ifelse(customers$PhoneCoTenure == 0, 1, customers$PhoneCoTenure)
customers$VoiceOverTenure <- ifelse(is.na(customers$VoiceOverTenure), customers$VoiceLastMonth, customers$VoiceOverTenure)
```

# Adding of additional variables 
```{r additionalVaribales, message=FALSE}
customers$LastMonthTotalValue <- customers$VoiceLastMonth + customers$EquipmentLastMonth + customers$DataLastMonth
customers$OverTenureTotalValue <- customers$VoiceOverTenure + customers$EquipmentOverTenure + customers$DataOverTenure
customers$TotalDebt <- customers$CreditDebt + customers$OtherDebt
customers$AverageMonthlyRevenue <- customers$OverTenureTotalValue/customers$PhoneCoTenure

customers$Gender_num <- ifelse(customers$Gender == "Male", 1, 2)
customers$LoanDefault_num <- ifelse(customers$LoanDefault == "Yes", 1, 2)
customers$MaritalStatus_num <- ifelse(customers$MaritalStatus == "Married", 1, 2)
customers$Multiline_num <- ifelse(customers$Multiline == "Yes", 1, 2)

customers <- mutate(customers, HHIncome_group = case_when(
 HHIncome %in%  0:25000          ~ 1,
 HHIncome %in%  25001:50000      ~ 2,
 HHIncome %in%  50001:75000      ~ 3,
 HHIncome %in%  75001:100000     ~ 4,
 HHIncome %in%  100001:150001    ~ 5,
 HHIncome >     150001           ~ 6
 )
) 
customers$DebtToIncomeRatio_int <- as.integer(10*(customers$DebtToIncomeRatio))
customers <- mutate(customers, DebtToIncomeRatio_group = case_when(
DebtToIncomeRatio_int %in%  0:50         ~ 5,
DebtToIncomeRatio_int %in%  51:100       ~ 10,
DebtToIncomeRatio_int %in%  101:150      ~ 15,
DebtToIncomeRatio_int %in%  151:200      ~ 20,
DebtToIncomeRatio_int %in%  201:250      ~ 25,
DebtToIncomeRatio_int %in%  251:300      ~ 30,
DebtToIncomeRatio_int %in%  301:400      ~ 40,
DebtToIncomeRatio_int >     401          ~ -1
 )
)  

customers <- mutate(customers, EducationYears_group = case_when(
 EducationYears %in%  6:10      ~ 1,
 EducationYears %in%  11:15     ~ 2,
 EducationYears %in%  16:19     ~ 3,
 EducationYears %in%  20:21     ~ 4,
 EducationYears %in%  22:23     ~ 5,
 EducationYears >     23        ~ 6
 )
)

customers <- mutate(customers, Age_group = case_when(
 Age %in%  0:18      ~ 1,
 Age %in%  18:40     ~ 2,
 Age %in%  40:60     ~ 3,
 Age >     60        ~ 4
 )
)

customers <- mutate(customers, EmploymentLength_group = case_when(
 EmploymentLength %in%  0:5       ~ 1,
 EmploymentLength %in%  6:10      ~ 2,
 EmploymentLength %in%  11:20     ~ 3,
 EmploymentLength %in%  21:30     ~ 4,
 EmploymentLength >     30        ~ 5
 )
)
```

# Create a dataset for analysis with required variables only
```{r filteredColumnData, message=FALSE}
customers.for.analysis <- customers[,c("Gender_num","Age","EducationYears","EmploymentLength",
                                     "HHIncome_group","DebtToIncomeRatio_group","TotalDebt",
                                     "MaritalStatus_num","HouseholdSize","HomeOwner",
                                     "PhoneCoTenure", "Multiline_num",
                                     "VoiceOverTenure","EquipmentOverTenure", "DataOverTenure",
                                     "OverTenureTotalValue","AverageMonthlyRevenue")]

#customers.for.analysis <- customers[,c("HHIncome_group","DebtToIncomeRatio_group",
#                                       "OverTenureTotalValue")]

customers.for.analysis <- customers[,c("Gender_num","Age_group","EducationYears_group", "EmploymentLength_group",
                                     "HHIncome_group","DebtToIncomeRatio_group", 
                                     "MaritalStatus_num","HouseholdSize","HomeOwner",
                                     "PhoneCoTenure", "Multiline_num",
                                     "OverTenureTotalValue","AverageMonthlyRevenue")]

str(customers.for.analysis)

summary(customers.for.analysis)
```

#Clustering analysis
```{r}
#summary(customers.for.analysis)
#sapply(customers.for.analysis, class)
#sum(is.na(customers.for.analysis))
#sapply(customers.for.analysis, function(x) sum(is.na(x))/length(x))*100

#Rescale the data before input into kmeans for creating clusters
customers.for.analysis_2 <- as.data.frame(scale(customers.for.analysis))
#class(cust_data.df.num_2)

customers.for.analysis_2.kmeans_seg_5 <- kmeans(customers.for.analysis_2, centers = 4, nstart = 25)
names(customers.for.analysis_2.kmeans_seg_5)
sol_5 <-fviz_cluster(customers.for.analysis_2.kmeans_seg_5, cust_data.df.num_2)
sol_5
fviz_cluster(k4, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 4")


k2 <- kmeans(customers.for.analysis_2, centers = 2, nstart = 25)
k3 <- kmeans(customers.for.analysis_2, centers = 3, nstart = 25)
k4 <- kmeans(customers.for.analysis_2, centers = 4, nstart = 25)
k5 <- kmeans(customers.for.analysis_2, centers = 5, nstart = 25)
#k6 <- kmeans(customers.for.analysis_2, centers = 6, nstart = 25)
#k7 <- kmeans(customers.for.analysis_2, centers = 7, nstart = 25)
#k8 <- kmeans(customers.for.analysis_2, centers = 8, nstart = 25)


# plots to compare
p2 <- fviz_cluster(k2, geom = "point", data = customers.for.analysis_2) + ggtitle("k = 2")
p3 <- fviz_cluster(k3, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 3")
p4 <- fviz_cluster(k4, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 4")
p5 <- fviz_cluster(k5, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 5")
#p6 <- fviz_cluster(k6, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 6")
#p7 <- fviz_cluster(k7, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 7")
#p8 <- fviz_cluster(k8, geom = "point",  data = customers.for.analysis_2) + ggtitle("k = 8")

library(gridExtra)
grid.arrange(p2, p3, p4, p5,  nrow = 2)
#grid.arrange(p6, p7, p8,  nrow = 2)


```


```{r elbowMethod}
set.seed(1234)
#Function to compute total within cluster sum of squares 
wss <- function(k) {
  kmeans(customers.for.analysis_2, k, nstart = 10)$tot.withinss
}

#Compute and plot the within sum of squares (wss) for k = 1 to k = 10
k.values <- 1:15

#Extract wss for 2 - 10 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
     type = "b", pch = 19, frame = FALSE,
     xlab = "Number of clusters K",
     ylab = "Total within clusters sum of squares")

set.seed(1234)
fviz_nbclust(customers.for.analysis_2, kmeans, method = "wss")
```

```{r silhouette}
avg_sil <- function(k) {
  km.res <- kmeans(customers.for.analysis_2, centers = k, nstart = 25)
  ss <- silhouette(km.res$cluster, dist(customers.for.analysis_2))
  mean(ss[, 3])
}

#Compute and plot wss for k = 2 to k = 10
k.values <-  2: 10

#Visually check the distribution of income according to segment
#box
#Extract average silhouette for 2 - 10 clusters
avg_sil_values <-  map_dbl(k.values, avg_sil)

# Based on the plots - does the 7 segment solution look optimal??
plot(k.values, avg_sil_values,
     type = "b", pch = 19, frame = FALSE,
     xlab = "Number of clusters K",
     ylab = "Total within clusters sum of squares")
                           
# Similar to the elbow method, this process to compute the “average silhoutte method” 
# has been wrapped up in a single function (fviz_nbclust):

fviz_nbclust(customers.for.analysis_2, kmeans, method = "silhouette")
```

```{r gapStat}
set.seed(123)
gap_stat <- clusGap(customers.for.analysis_2, FUN = kmeans, nstart = 25,
                  K.max = 10, B = 10)
# Print the result
print(gap_stat, method = "firstmax")
fviz_gap_stat(gap_stat)
```