---
title: "Data-Due-Diligence"
author: "Amol Gote"
date: "6/6/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(maps)
library(scales)
library(sf)
library(Hmisc)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(rpart)
library(rpart.plot)
library(poLCA)
library(dendextend)
library(gridExtra)
```

```{r include = FALSE}
#knitr::opts_chunk$set(echo=FALSE)
```

```{r loadData, message=FALSE}
customers <- read_csv("data/Customer_Dataset_File_Original.csv")
```
# Feature engineering steps
# Data Imputation for customers.
```{r imputeGender, message=FALSE}
i <- 0
count <- 0
for(i in 1:nrow(customers)){
  if(is.na(customers$Gender[i])){
    count <- count + 1
    if (count %% 2 == 0){
      customers$Gender[i] <- "Female"
    }
    else{
      customers$Gender[i] <- "Male"
    }
  }
}
```

# Data Imputation for HouseholdSize.
```{r imputeHouseHold, message=FALSE}
customers$HouseholdSize[is.na(customers$HouseholdSize)] <- median(customers$HouseholdSize, na.rm = T)
```

# Data Imputation for HomeOwner.
```{r imputeHoeOwnership, message=FALSE}
get_mode <- function(x) {
  unique_x <- unique(x)
  mode <- unique_x[which.max(tabulate(match(x, unique_x)))]
  mode
}
mode_value = get_mode(customers$HomeOwner)
customers$HomeOwner[is.na(customers$HomeOwner)] <- mode_value
```

# Data Imputation for JobCategory
```{r imputeJobCategory, message=FALSE}
customers$JobCategory[is.na(customers$JobCategory)] <- "Misc"
```

# Remove the dollar sign from household income
```{r cleanAmountVariables, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
customers <- customers %>%
  replace(.=="#NULL!", NA) # replace with NA
customers$HHIncome = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$HHIncome)))) 
customers$VoiceLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$VoiceLastMonth)))) 
customers$VoiceOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$VoiceOverTenure)))) 
customers$CardSpendMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$CardSpendMonth)))) 
customers$EquipmentLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$EquipmentLastMonth)))) 
customers$EquipmentOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$EquipmentOverTenure)))) 
customers$DataLastMonth = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$DataLastMonth)))) 
customers$DataOverTenure = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$DataOverTenure)))) 
customers$CardSpendMonth[is.na(customers$CardSpendMonth)] <- 0
sapply(customers,function(x)sum(is.na(x)))
```


# Remove the dollar sign from CarValue
```{r imputeCarValue, message=FALSE}
#customers <- read_csv("data/Customer_Dataset_File_Original.csv")
customers$CarValue = gsub("\\ ", "", customers$CarValue) 
summary(customers$CarValue)
carValueLessThanZero <- customers %>% 
  filter(CarValue =="$(1,000.00)") %>% 
  nrow()
customers$CarValue[customers$CarValue =="$(1,000.00)"] <- 0
customers$CarValue = as.numeric(gsub("\\,", "", gsub("\\$", "", gsub("\\-", "", customers$CarValue)))) 
customers$CarValue = as.numeric(customers$CarValue)
carValueLessThanZero <- customers %>% 
  filter(is.na(customers$CarValue)) %>%
  nrow()
  
carOwnership <- customers %>% 
  filter(customers$CarOwnership == "-1") %>%
  nrow()
carBrand <- customers %>% 
  filter(customers$CarBrand == "-1") %>%
  nrow()
customers$CarOwnership[customers$CarOwnership =="-1"] <- NA
customers$CarBrand[customers$CarBrand =="-1"] <- NA
```

# Impute Commute Time, there are 2 missing values
```{r imputeCommuteTime, message=FALSE}
customers$CommuteTime = as.numeric(customers$CommuteTime)
customers$CommuteTime[is.na(customers$CommuteTime)] <- mean(customers$CommuteTime, na.rm = T)
```
#Address Missing values for EquipmentLastMonth, DataLastMonth, EquipmentOverTenure,DataOverTenure
```{r addressMissingValies, message=FALSE}
customers$EquipmentLastMonth[is.na(customers$EquipmentLastMonth)] <- 0
customers$DataLastMonth[is.na(customers$DataLastMonth)] <- 0
customers$EquipmentOverTenure[is.na(customers$EquipmentOverTenure)] <- 0
customers$DataOverTenure[is.na(customers$DataOverTenure)] <- 0
customers$PhoneCoTenure <- ifelse(customers$PhoneCoTenure == 0, 1, customers$PhoneCoTenure)
customers$VoiceOverTenure <- ifelse(is.na(customers$VoiceOverTenure), customers$VoiceLastMonth, customers$VoiceOverTenure)
```

# Adding of additional variables 
```{r additionalVariables, message=FALSE}
customers$LastMonthTotalValue <- customers$VoiceLastMonth + customers$EquipmentLastMonth + customers$DataLastMonth
customers$OverTenureTotalValue <- customers$VoiceOverTenure + customers$EquipmentOverTenure + customers$DataOverTenure
customers$TotalDebt <- customers$CreditDebt + customers$OtherDebt
customers$AverageMonthlyRevenue <- customers$OverTenureTotalValue/customers$PhoneCoTenure

customers$Gender_num <- ifelse(customers$Gender == "Male", 1, 2)
customers$LoanDefault_num <- ifelse(customers$LoanDefault == "Yes", 1, 2)
customers$MaritalStatus_num <- ifelse(customers$MaritalStatus == "Married", 1, 2)
customers$Multiline_num <- ifelse(customers$Multiline == "Yes", 1, 2)

customers <- mutate(customers, HHIncome_group = case_when(
 HHIncome %in%  0:25000          ~ 1,
 HHIncome %in%  25001:50000      ~ 2,
 HHIncome %in%  50001:75000      ~ 3,
 HHIncome %in%  75001:100000     ~ 4,
 HHIncome %in%  100001:150001    ~ 5,
 HHIncome >     150001           ~ 6
 )
) 
customers$DebtToIncomeRatio_int <- as.integer(10*(customers$DebtToIncomeRatio))
customers <- mutate(customers, DebtToIncomeRatio_group = case_when(
DebtToIncomeRatio_int %in%  0:50         ~ 5,
DebtToIncomeRatio_int %in%  51:100       ~ 10,
DebtToIncomeRatio_int %in%  101:150      ~ 15,
DebtToIncomeRatio_int %in%  151:200      ~ 20,
DebtToIncomeRatio_int %in%  201:250      ~ 25,
DebtToIncomeRatio_int %in%  251:300      ~ 30,
DebtToIncomeRatio_int %in%  301:400      ~ 40,
DebtToIncomeRatio_int >     401          ~ 100
 )
)  

customers <- mutate(customers, EducationYears_group = case_when(
 EducationYears %in%  6:10      ~ 1,
 EducationYears %in%  11:15     ~ 2,
 EducationYears %in%  16:19     ~ 3,
 EducationYears %in%  20:21     ~ 4,
 EducationYears %in%  22:23     ~ 5,
 EducationYears >     23        ~ 6
 )
)

customers <- mutate(customers, EducationYears_category = case_when(
 EducationYears %in%  6:10      ~ '6-10 years',
 EducationYears %in%  11:15     ~ '11-15 years',
 EducationYears %in%  16:19     ~ '16-19 years',
 EducationYears %in%  20:21     ~ '20-21 years',
 EducationYears %in%  22:23     ~ '22-23 years',
 EducationYears >     23        ~ '> 23 years'
 )
)  

customers <- mutate(customers, Age_group = case_when(
 Age %in%  0:18      ~ 1,
 Age %in%  18:40     ~ 2,
 Age %in%  40:60     ~ 3,
 Age >     60        ~ 4
 )
)

customers <- mutate(customers, Age_category = case_when(
 Age %in%  0:18      ~ "Below 18",
 Age %in%  18:40     ~ "18 - 40",
 Age %in%  40:60     ~ "40 - 60",
 Age >     60        ~ "> 60"
 )
)

customers <- mutate(customers, EmploymentLength_group = case_when(
 EmploymentLength %in%  0:5       ~ 1,
 EmploymentLength %in%  6:10      ~ 2,
 EmploymentLength %in%  11:20     ~ 3,
 EmploymentLength %in%  21:30     ~ 4,
 EmploymentLength >     30        ~ 5
 )
)
```

# Create a dataset for analysis with required variables only
```{r filteredColumnData, message=FALSE}
customers.for.analysis <- customers[,c("Gender_num","Age","EducationYears","EmploymentLength",
                                     "HHIncome_group","DebtToIncomeRatio_group","TotalDebt",
                                     "MaritalStatus_num","HouseholdSize","HomeOwner",
                                     "PhoneCoTenure", "Multiline_num",
                                     "VoiceOverTenure","EquipmentOverTenure", "DataOverTenure",
                                     "OverTenureTotalValue","AverageMonthlyRevenue")]

customers.for.analysis <- customers[,c("Gender_num","Age_group","EducationYears_group", "EmploymentLength_group",
                                     "HHIncome_group","DebtToIncomeRatio_group", 
                                     "MaritalStatus_num","HouseholdSize","HomeOwner",
                                     "PhoneCoTenure", "Multiline_num",
                                     "OverTenureTotalValue","AverageMonthlyRevenue")]

str(customers.for.analysis)

summary(customers.for.analysis)
```

#Clustering analysis
```{r kMeanClusterAnalysis, message=FALSE}
#Rescale the data before input into kmeans for creating clusters
customers.for.analysis_2 <- as.data.frame(scale(customers.for.analysis))

customers.for.analysis_2.kmeans_seg_5 <- kmeans(customers.for.analysis_2, centers = 4, nstart = 25)
names(customers.for.analysis_2.kmeans_seg_5)

k2 <- kmeans(customers.for.analysis_2, centers = 2, nstart = 25)
k3 <- kmeans(customers.for.analysis_2, centers = 3, nstart = 25)
k4 <- kmeans(customers.for.analysis_2, centers = 4, nstart = 25)
k5 <- kmeans(customers.for.analysis_2, centers = 5, nstart = 25)
k6 <- kmeans(customers.for.analysis_2, centers = 6, nstart = 25)
k7 <- kmeans(customers.for.analysis_2, centers = 7, nstart = 25)
k8 <- kmeans(customers.for.analysis_2, centers = 8, nstart = 25)
k9 <- kmeans(customers.for.analysis_2, centers = 9, nstart = 25)


# plots to compare
p2 <- fviz_cluster(k2, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 2")
p3 <- fviz_cluster(k3, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 3")
p4 <- fviz_cluster(k4, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 4")
p5 <- fviz_cluster(k5, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 5")
p6 <- fviz_cluster(k6, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 6")
p7 <- fviz_cluster(k7, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 7")
p8 <- fviz_cluster(k8, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 8")
p9 <- fviz_cluster(k9, geom = "point",  data = customers.for.analysis_2, ggtheme = theme_minimal()) + ggtitle("k = 9")


grid.arrange(p2, p3, p4, p5,  nrow = 2)
grid.arrange(p6, p7, p8, p9,  nrow = 2)
p4
```


```{r kMeanElbowMethod, message=FALSE}
set.seed(1234)
#Function to compute total within cluster sum of squares 
wss <- function(k) {
  kmeans(customers.for.analysis_2, k, nstart = 25, iter.max = 30)$tot.withinss
}

#Compute and plot the within sum of squares (wss) for k = 1 to k = 10
k.values <- 1:15

#Extract wss for 2 - 10 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
     type = "b", pch = 19, frame = FALSE,
     xlab = "Number of clusters K",
     ylab = "Total within clusters sum of squares")

set.seed(1234)
fviz_nbclust(customers.for.analysis_2, kmeans, method = "wss")
```

```{r lMeansilhouette, message=FALSE}
avg_sil <- function(k) {
  km.res <- kmeans(customers.for.analysis_2, centers = k, nstart = 25, iter.max = 50)
  ss <- silhouette(km.res$cluster, dist(customers.for.analysis_2))
  mean(ss[, 3])
}

#Compute and plot wss for k = 2 to k = 10
k.values <-  2: 10

#Visually check the distribution of income according to segment
#box
#Extract average silhouette for 2 - 10 clusters
avg_sil_values <-  map_dbl(k.values, avg_sil)

# Based on the plots - does the 7 segment solution look optimal??
plot(k.values, avg_sil_values,
     type = "b", pch = 19, frame = FALSE,
     xlab = "Number of clusters K",
     ylab = "Total within clusters sum of squares")
                           
# Similar to the elbow method, this process to compute the “average silhoutte method” 
# has been wrapped up in a single function (fviz_nbclust):

fviz_nbclust(customers.for.analysis_2, kmeans, method = "silhouette")
```

```{r kMeangapStat}
#set.seed(123)
#gap_stat <- clusGap(customers.for.analysis_2, FUN = kmeans, nstart = 25,
                 # K.max = 10, B = 10)
# Print the result
#print(gap_stat, method = "firstmax")
#fviz_gap_stat(gap_stat)
```

```{r decisionTree}
customers.for.analysis_tree <- customers[,c("Gender","Age","EducationYears", "EmploymentLength",
                                     "HHIncome","DebtToIncomeRatio", 
                                     "MaritalStatus","HouseholdSize","HomeOwner",
                                     "PhoneCoTenure", "Multiline",
                                     "OverTenureTotalValue")]

tree <- rpart(OverTenureTotalValue~., 
              customers.for.analysis_tree)
tree
rpart.plot(tree, box.palette="Blues", extra = 101)
```
```{r latentClassAnalysis}
f <- cbind(Gender_num,Age_group,EducationYears_group,EducationYears_group,EmploymentLength_group,HHIncome_group,
           MaritalStatus_num)~1
poLCA(f, customers.for.analysis, nclass=3, nrep=5, na.rm=FALSE, verbose = FALSE)
```

```{r HierarchicalClusterAnalysis }
customers.for.analysis_hierarchy <- customers[,c("Gender_num","Age_group","EducationYears_group", "EmploymentLength_group",
                                     "HHIncome_group","DebtToIncomeRatio_group", 
                                     "MaritalStatus_num","HouseholdSize","HomeOwner",
                                     "PhoneCoTenure", "Multiline_num",
                                     "OverTenureTotalValue","AverageMonthlyRevenue")]

# compute divisive hierarchical clustering
hc4 <- diana(customers.for.analysis_hierarchy)

# Divise coefficient; amount of clustering structure found
#hc4$dc
## [1] 0.8514345

# plot dendrogram
#pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of diana")
```

```{r HierarchicalClusterAnalysisPlots, message=FALSE}
hc4$dc
pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of diana")
clust2 <- cutree(hc4, k = 2)
p2 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust2), ggtheme = theme_minimal()) + ggtitle("k = 2")

clust3 <- cutree(hc4, k = 3)
p3 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust3), ggtheme = theme_minimal()) + ggtitle("k = 3")  

clust4 <- cutree(hc4, k = 4)
p4 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust4), ggtheme = theme_minimal()) + ggtitle("k = 4")  

clust5 <- cutree(hc4, k = 5)
p5 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust5), ggtheme = theme_minimal()) + ggtitle("k = 5")  

clust6 <- cutree(hc4, k = 6)
p6 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust6), ggtheme = theme_minimal()) + ggtitle("k = 6")

clust7 <- cutree(hc4, k = 7)
p7 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust7), ggtheme = theme_minimal()) + ggtitle("k = 7")  

clust8 <- cutree(hc4, k = 8)
p8 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust8), ggtheme = theme_minimal()) + ggtitle("k = 8")

clust9 <- cutree(hc4, k = 9)
p9 <- fviz_cluster(list(data = customers.for.analysis_hierarchy, cluster = clust9), ggtheme = theme_minimal()) + ggtitle("k = 9")


grid.arrange(p2, p3, p4, p5,  nrow = 2)
grid.arrange(p6, p7, p8, p9,  nrow = 2)
p4
```
```{r HierarchicalclusterVerification, message=FALSE}
fviz_nbclust(customers.for.analysis_hierarchy, FUN = hcut, method = "wss")
fviz_nbclust(customers.for.analysis_hierarchy, FUN = hcut, method = "silhouette")
```

```{r HierarchicalclusterSummary, message=FALSE}
summary(k4)
```

```{r clusterDataAnalysis, message=FALSE}
customers$cluster <- k4$cluster
```

```{r analysisBasedonCluster}
groupedData <- customers %>% 
     group_by(cluster) %>% 
     summarise(totalValue= sum(OverTenureTotalValue)) 
groupedData

ggplot(data=groupedData, aes(x=cluster, y=totalValue)) +
  geom_bar(colour="black", stat="identity")

groupedData <- customers %>% 
     group_by(cluster) %>% 
     summarise(averageMonthlyRevenue= mean(AverageMonthlyRevenue)) 

ggplot(data=groupedData, aes(x=cluster, y=averageMonthlyRevenue)) +
  geom_bar(colour="black", stat="identity")

ggplot(customers, aes(cluster)) +
  geom_bar(fill = "#0073C2FF") 

ggplot(data = customers, aes(x = cluster, y = OverTenureTotalValue, group=cluster)) +
  geom_boxplot() +
  labs( title= "Distribution of monthly revenue generated by job category", x="Job category", y = "Monthly revenue ($)")  +
  theme_minimal() 

ggplot(data = customers, aes(x = cluster, y = AverageMonthlyRevenue, group=cluster)) +
  geom_boxplot() +
  labs( title= "Distribution of monthly revenue generated by job category", x="Job category", y = "Monthly revenue ($)")  +
  theme_minimal() 

```

```{r assignCustomerCategories}
customers$category <- ""
for(i in 1:nrow(customers)){
    count <- count + 1
    if (customers$cluster[i] == 1){
        customers$category[i] <- "Platinum"
    }
    else if (customers$cluster[i] == 2){
      customers$category[i] <- "Bronze"
    }
    else if (customers$cluster[i] == 3){
      customers$category[i] <- "Gold"
    }
    else if (customers$cluster[i] == 4){
      customers$category[i] <- "Silver"
    }
}

platinum_customers <- customers %>%
  filter(category == "Platinum")

gold_customers <- customers %>%
  filter(category == "Gold")

silver_customers <- customers %>%
  filter(category == "Silver")

bronze_customers <- customers %>%
  filter(category == "Bronze")

```


```{r attributeComparison_HHIncom}
summary(platinum_customers$HHIncome)
summary(gold_customers$HHIncome)
summary(silver_customers$HHIncome)
summary(bronze_customers$HHIncome)
```
```{r attributeComparison_Householdsize}
summary(platinum_customers$HouseholdSize)
summary(gold_customers$HouseholdSize)
summary(silver_customers$HouseholdSize)
summary(bronze_customers$HouseholdSize)
```

```{r attributeComparison_PhoneCoTenure}
summary(platinum_customers$PhoneCoTenure)
summary(gold_customers$PhoneCoTenure)
summary(silver_customers$PhoneCoTenure)
summary(bronze_customers$PhoneCoTenure)
```

```{r attributeComparison_OverTenureTotalValue}
summary(platinum_customers$OverTenureTotalValue)
summary(gold_customers$OverTenureTotalValue)
summary(silver_customers$OverTenureTotalValue)
summary(bronze_customers$OverTenureTotalValue)
```

```{r attributeComparison_AverageMonthlyRevenue}
summary(platinum_customers$AverageMonthlyRevenue)
summary(gold_customers$AverageMonthlyRevenue)
summary(silver_customers$AverageMonthlyRevenue)
summary(bronze_customers$AverageMonthlyRevenue)
```

```{r attributeComparison_EducationYears}
summary(platinum_customers$EducationYears)
summary(gold_customers$EducationYears)
summary(silver_customers$EducationYears)
summary(bronze_customers$EducationYears)
```


```{r attributeComparison_Age}
summary(platinum_customers$Age)
summary(gold_customers$Age)
summary(silver_customers$Age)
summary(bronze_customers$Age)
```
```{r attributeComparison_DebtToIncomeRatio}
summary(platinum_customers$DebtToIncomeRatio)
summary(gold_customers$DebtToIncomeRatio)
summary(silver_customers$DebtToIncomeRatio)
summary(bronze_customers$Age)
```

```{r visualizations, message=FALSE}
ggplot(customers) +
    geom_bar(aes(x = category, fill = Gender))

ggplot(customers) +
    geom_bar(aes(x = category, fill = CarOwnership))

ggplot(customers) +
    geom_bar(aes(x = category, fill = Age_category))

ggplot(customers) +
    geom_bar(aes(x = category, fill = MaritalStatus))

ggplot(customers) +
    geom_bar(aes(x = category, fill = Multiline))


ggplot(customers, aes(x = HHIncome, y = AverageMonthlyRevenue)) + 
  geom_point(color = "steelblue") +
  facet_grid(~category) +
  geom_smooth() + 
  labs( title= "Monthly revenue generated Debt to income ratio", x="Debt to income ratio (DTI)", y = "Over Tenure Total Value ($)") +
  theme_minimal() 

```

```{r customerSegmentsByHHIncomeByAvgMonthlyValue, message=FALSE}
groupedData <- customers %>% 
      filter(HHIncome < 300000) %>%
     group_by(category,HHIncome) %>% 
     summarise(totalMonthlyValue= mean(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = HHIncome, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category, nrow(2)) +
  labs( title= "Average Monthly revenue generated to House hold income", x="Household Income ($)", y = "Average Monthly revenue ($)") +
  scale_x_continuous(labels = comma) +
  theme_minimal() 

```

```{r customerSegmentsByHHIncomeByTotalMonthlyValue, message=FALSE}
groupedData <- customers %>% 
      filter(HHIncome < 300000) %>%
     group_by(category,HHIncome) %>% 
     summarise(totalMonthlyValue= sum(OverTenureTotalValue)) 

ggplot(groupedData, aes(x = HHIncome, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category, nrow(2)) +
    scale_x_continuous(labels = comma) +
  theme_minimal() 

```

```{r variousParamtersDistributions, message=FALSE}

ggplot(customers) +
    geom_bar(aes(x = category, fill = EducationYears_category))


groupedData <- customers %>% 
  group_by(category) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  mutate(ypos = cumsum(per) - 0.5*per ) %>% 
  arrange(desc(category))

groupedData$label <- scales::percent(round(groupedData$per,2))

write_csv(groupedData,"data/cluster_distribution.csv")
ggplot(data=groupedData) +
  geom_bar(aes(x="", y=per, fill=category), stat="identity", width = 1) +
  coord_polar("y", start=0, direction = -1) +
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  theme_void()

ggplot(data=groupedData) +
  geom_bar(aes(x="", y=per, fill=category), stat="identity", width = 1) +
  coord_polar("y", start=0, direction = -1) +
  scale_fill_brewer(palette="Set1") +
  theme_void()


ggplot(groupedData, aes(x="", y=per, fill=category)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  theme(legend.position="none") +
  
  geom_text(aes(y = ypos, label = category), color = "white", size=6) +
  scale_fill_brewer(palette="Set1")

groupedData <- customers %>%
  group_by(category) %>%
  summarise(totalValue= sum(OverTenureTotalValue), categoryCount = n(), averageValue= mean(AverageMonthlyRevenue)) 
write_csv(groupedData,"data/cluster_total_revenue.csv")


groupedData <- customers %>%
  group_by(category) %>%
  summarise(averageValue= mean(Age)) 

groupedData
write_csv(groupedData,"data/cluster_total_revenue.csv")
```

```{r platinumMaritalStatusDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  group_by(MaritalStatus) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = MaritalStatus, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "steelblue") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r goldMaritalStatusDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Gold") %>%
  group_by(MaritalStatus) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = MaritalStatus, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "steelblue") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r silverMaritalStatusDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Silver") %>%
  group_by(MaritalStatus) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = MaritalStatus, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "steelblue") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r bronzeMaritalStatusDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Bronze") %>%
  group_by(MaritalStatus) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = MaritalStatus, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "steelblue") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r platinumGenderDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  group_by(Gender) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = Gender, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "DarkGreen") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r goldGenderDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Gold") %>%
  group_by(Gender) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = Gender, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "DarkGreen") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r silverGenderDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Silver") %>%
  group_by(Gender) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = Gender, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "DarkGreen") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```
```{r bronzeGenderDistributions, message=FALSE}

cystomersFiltered <- customers %>%
  filter(category == "Bronze") %>%
  group_by(Gender) %>%
  summarise(totalCount = n())

ggplot(cystomersFiltered, aes(x = Gender, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "DarkGreen") + 
  geom_text(aes(label=totalCount), vjust=-0.5) +
  scale_y_continuous() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    aspect.ratio = 2/1) 
```


```{r educationYearsAcrossCustomerSegments, message=FALSE}

customersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  group_by(EducationYears_category) %>%
  summarise(totalCount = n()) 

write_csv(customersFiltered,"data/Platinum_education.csv")

ggplot(customersFiltered, aes(x = EducationYears_category, y= totalCount, label = totalCount)) +
  geom_col(width = 0.2, fill = "DarkGreen") + 
  geom_text(aes(label=totalCount), hjust=-0.5) +
  scale_y_continuous() +
  coord_flip() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 
```


```{r multilineAcrossCustomerSegments, message=FALSE}

customersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  group_by(Multiline) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Platinum_Multiline.csv")

customersFiltered <- customers %>%
  filter(category == "Gold") %>%
  group_by(Multiline) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Gold_Multiline.csv")

customersFiltered <- customers %>%
  filter(category == "Silver") %>%
  group_by(Multiline) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Silver_Multiline.csv")

customersFiltered <- customers %>%
  filter(category == "Bronze") %>%
  group_by(Multiline) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Bronze_Multiline.csv")

```

```{r loanDefaultsAcrossSegments, message=FALSE}

customersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  group_by(LoanDefault) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Platinum_LoanDefault.csv")

customersFiltered <- customers %>%
  filter(category == "Gold") %>%
  group_by(LoanDefault) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Gold_LoanDefault.csv")

customersFiltered <- customers %>%
  filter(category == "Silver") %>%
  group_by(LoanDefault) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Silver_LoanDefault.csv")

customersFiltered <- customers %>%
  filter(category == "Bronze") %>%
  group_by(LoanDefault) %>%
  summarise(totalCount = n()) 
write_csv(customersFiltered,"data/Bronze_LoanDefault.csv")


```

```{r carOwnershipAcrossSegments, message=FALSE}

customersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  drop_na() %>% 
  group_by(CarOwnership) %>%
  summarise(totalCount = n())

write_csv(customersFiltered,"data/Platinum_Car.csv")

customersFiltered <- customers %>%
  filter(category == "Gold") %>%
  drop_na() %>% 
  group_by(CarOwnership) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Gold_Car.csv")

customersFiltered <- customers %>%
  filter(category == "Silver") %>%
  drop_na() %>% 
  group_by(CarOwnership) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Silver_Car.csv")

customersFiltered <- customers %>%
  filter(category == "Bronze") %>%
  drop_na() %>% 
  group_by(CarOwnership) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Bronze_Car.csv")

```

```{r variousRevenueStreamsAcrossSegments, message=FALSE}
customersFiltered <- customers %>%
  group_by(category) %>%
  summarise(voiceRevenue = sum(VoiceOverTenure), equipmentRevenue = sum(EquipmentOverTenure), dataRevenue = sum(DataOverTenure))
write_csv(customersFiltered,"data/revenue_voice_equip_data.csv")

customersFiltered <- customers %>%
  drop_na() %>% 
  group_by(category, EquipmentRental) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Count_EquipmentRental.csv")

customersFiltered <- customers %>%
  drop_na() %>% 
  group_by(category, WirelessData) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Count_WirelessData.csv")

```


```{r customerAcountAcrossSegmentss, message=FALSE}

customers$hierarchicalCluster <- clust4
customersGroupedData <- customers %>%
  group_by(hierarchicalCluster) %>%
  summarise(totalCount = n())

customersGroupedData
```


```{r eBillingAcrosssegments, message=FALSE}
customersFiltered <- customers %>%
  filter(category == "Platinum") %>%
  drop_na() %>% 
  group_by(EBilling) %>%
  summarise(totalCount = n())

write_csv(customersFiltered,"data/Platinum_EBilling.csv")

customersFiltered <- customers %>%
  filter(category == "Gold") %>%
  drop_na() %>% 
  group_by(EBilling) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Gold_EBilling.csv")

customersFiltered <- customers %>%
  filter(category == "Silver") %>%
  drop_na() %>% 
  group_by(EBilling) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Silver_EBilling.csv")

customersFiltered <- customers %>%
  filter(category == "Bronze") %>%
  drop_na() %>% 
  group_by(EBilling) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/Bronze_EBilling.csv")
```

```{r goldNoMultiLineMarriedCustomers, message=FALSE}
customersFiltered <- customers %>%
  filter(category == "Gold") %>%
  filter(Multiline == "No") %>%
  filter(MaritalStatus == "Married") %>%
  drop_na() %>% 
  group_by(category) %>%
  summarise(totalCount = n())
customersFiltered
```

```{r silverNoMultiLineMarriedCustomers, message=FALSE}
customersFiltered <- customers %>%
  filter(category == "Silver") %>%
  filter(Multiline == "No") %>%
  filter(MaritalStatus == "Married") %>%
  drop_na() %>% 
  group_by(category) %>%
  summarise(totalCount = n())
customersFiltered
```


```{r carsOwnerAcrossSegments, message=FALSE}
customersFiltered <- customers %>%
  group_by(category) %>%
  summarise(totalCountNUmberOfCars = sum(CarsOwned), numberOfCustomer = n())
customersFiltered$carRatio = customersFiltered$totalCountNUmberOfCars/customersFiltered$numberOfCustomer
customersFiltered
```


```{r jobCategoryCountAcrossSegments, message=FALSE}
customersFiltered <- customers %>%
  group_by(category, JobCategory) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/all_jogbcategory.csv")
```

```{r creditCardCount, message=FALSE}
customersFiltered <- customers %>%
  group_by(category, CreditCard) %>%
  summarise(totalCount = n())
write_csv(customersFiltered,"data/all_creditcard.csv")
```

```{r creditCardSpendAcrossSegments, message=FALSE}
filteredCustomer <- customers %>%
  filter(CardSpendMonth <= 20000)

ggplot(filteredCustomer, aes(x=category, y=CardSpendMonth)) +
  geom_boxplot() +
  labs( title= "Credit Card Spend across segments", x="Customer Segments", y = "Credit Card Spend ($)") +
  theme_minimal() 

```

#Statistical analysis for Debt to income ratio
```{r averageMonthlyRevenueByHHIncome, message=FALSE}
lbls <- c('0-10','10-20','20-30', '30-40','40-50','50-60','60-70', '70-80', '80-90', '90-100')
groupedData <- customers %>% 
     group_by(category, DebtToIncomeRatio) %>% 
     summarise(totalMonthlyValue= sum(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = DebtToIncomeRatio, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth() + 
  facet_grid(~category) +
  labs( title= "Monthly revenue generated Debt to income ratio", x="Debt to income ratio (DTI)", y = "Monthly revenue ($)") +
  theme_minimal() 


groupedData <- customers %>% 
  filter(HHIncome <= 300000) %>%
  group_by(category, HHIncome) %>% 
  summarise(totalMonthlyValue= mean(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = HHIncome, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category) +
  labs( title= "Average Monthly revenue generated to House hold income", x="Household Income ($)", y = "Average Monthly revenue ($)") +
  scale_x_continuous(labels = comma) +
  theme_minimal() 


ggplot(groupedData, aes(x = HHIncome, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category) +
  labs( title= "Average Monthly revenue generated to House hold income", x="Household Income ($)", y = "Average Monthly revenue ($)") +
  scale_x_continuous(labels = comma) +
  theme_minimal() 

```


```{r avgMonthlyRevenueByCreditCardSpend, message=FALSE}
groupedData <- customers %>% 
  filter(HHIncome <= 300000) %>%
  group_by(category, CardSpendMonth) %>% 
  summarise(totalMonthlyValue= mean(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = CardSpendMonth, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 


```

```{r PhoneCoTenureByCardTenure, message=FALSE}
ggplot(customers, aes(x = PhoneCoTenure, y = CardTenure)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 
```


```{r avgMonthlyByCreditCardTenure, message=FALSE}
groupedData <- customers %>% 
  filter(HHIncome <= 300000) %>%
  group_by(category, CardTenure) %>% 
  summarise(totalMonthlyValue= mean(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = CardTenure, y = totalMonthlyValue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 
```


```{r avgMonthlyByHouseholdSize, message=FALSE}
groupedData <- customers %>% 
  group_by(category, HouseholdSize) %>% 
  summarise(totalMonthlyValue= mean(AverageMonthlyRevenue)) 

ggplot(customers, aes(x = HouseholdSize, y = AverageMonthlyRevenue)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 
```

```{r avgVoiceOverTenure, message=FALSE}

groupedData <- customers %>% 
  group_by(category, PhoneCoTenure) %>% 
  summarise(avgVoiceOverTenure= mean(VoiceOverTenure)) 

ggplot(groupedData, aes(x = PhoneCoTenure, y = avgVoiceOverTenure)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  labs( title= "Average voice revenue ($) by tenure", x="Tenure", y = "Average voice revenue ($)") +
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 
```


```{r avgEquipmentOverTenure, message=FALSE}

groupedData <- customers %>% 
  group_by(category, PhoneCoTenure) %>% 
  summarise(avgEquipmentOverTenure= mean(EquipmentOverTenure)) 

ggplot(groupedData, aes(x = PhoneCoTenure, y = avgEquipmentOverTenure)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  labs( title= "Average equipment revenue ($) by tenure", x="Tenure", y = "Average equipment revenue ($)") +
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 
```

```{r avgDataOverTenure, message=FALSE}

groupedData <- customers %>% 
  group_by(category, PhoneCoTenure) %>% 
  summarise(avgDataOverTenure= mean(DataOverTenure)) 

ggplot(groupedData, aes(x = PhoneCoTenure, y = avgDataOverTenure)) + 
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) + 
  labs( title= "Average data revenue ($) by tenure", x="Tenure", y = "Average data revenue ($)") +
  facet_wrap(~category) +
  scale_x_continuous(labels = comma) +
  theme_minimal() 
```

```{r averageMonthlyRevenueByPhoneCoTenureLine, message=FALSE}

groupedData <- customers %>% 
  filter(category != 'Platinum') %>%
  group_by(category, PhoneCoTenure) %>% 
  summarise(avgOverTenure= mean(AverageMonthlyRevenue)) 

ggplot(groupedData, aes(x = PhoneCoTenure, y = avgOverTenure, group = category, colour=category)) + 
  geom_line() +
  scale_x_continuous(labels = comma) +
  theme_minimal() 

```

```{r densityPlotsForAvgMonthRevenures, message=FALSE}

filteredCustomers <- customers %>% 
  filter(category != 'Platinum')

ggplot(filteredCustomers, aes(x=AverageMonthlyRevenue, fill=category)) + 
  geom_density(alpha=0.5) +
  labs( title= "Density plot for average monthly revenue", x="Average monthly revenue ($)", y = "density") +
  theme_minimal() 

filteredCustomers <- customers %>% 
  filter(category == 'Platinum')

ggplot(filteredCustomers, aes(x=AverageMonthlyRevenue, fill=category)) + 
  geom_density(alpha=.5) +
  labs( title= "Density plot for average monthly revenue", x="Average monthly revenue ($)", y = "density") +
  theme_minimal() 
```
